### 2022-11-25 paper2.tex
paper 2 start

\section*{Abstract}
Replication studies are increasingly conducted in order to confirm original
findings. However, there is no established standard how to assess replication
success and in practice many different approaches are used. The purpose of this
paper is to refine and extend a recently proposed reverse-Bayes approach for the
analysis of replication studies. We show how this method is directly related to
the relative effect size, the ratio of the replication to the original effect
estimate. This perspective leads to a new proposal to recalibrate the assessment
of replication success, the golden level. The recalibration ensures that for
borderline significant original studies replication success can only be achieved
if the replication effect estimate is larger than the original one. Conditional
power for replication success can then take any desired value if the original
study is significant and the replication sample size is large enough. Compared
to the standard approach to require statistical significance of both the
original and replication study, replication success at the golden level offers
uniform gains in project power and controls the Type-I error rate if the
replication sample size is not smaller than the original one. An application to
data from four large replication projects shows that the new approach leads to
more appropriate inferences, as it penalizes shrinkage of the replication
estimate compared to the original one, while ensuring that both effect estimates
are sufficiently convincing on their own.

\textbf{Key words}: Power, replication studies, sceptical $p$-value, shrinkage,
two-trials rule, type-I error rate

\section{Introduction}
Replication studies are conducted in order to investigate whether an original
finding can be confirmed in an independent study. Although replication has long
been a central part of the scientific method in many fields, the so-called
replication crisis \citep{Ioannidis2005, Begley2015} has led to increased
interest in replication over the last decade. These developments eventually
culminated in large-scale replication projects that were conducted in various
fields \citep{Errington2014, Klein2014, OSC2015, Ebersole2016, Camerer2016,
  Camerer2018, Cova2018, Klein2018}.

% Declaring a replication as successful is, however, not a
% straightforward task, and currently used approaches include
% \emph{statistical} significance of both the original and replication
% study,
%  compatibility of their effect estimates, and meta-analysis of
% the effect estimates. \emph{Many of the replication projects listed above also
%   report the relative effect size, the ratio of the
%   replication to the original effect estimate. For example, in \citet{Camerer2018}
%   the replication effect estimates were only half as large as the original ones
%   on average and even smaller in \citet{OSC2015}. This gives
%   clear evidence of a systematic bias of the original studies and
%   strongly suggests that the original and replication study should not
%   be treated as exchangeable. However, all the approaches mentioned
%   above will give the same results if the order of studies would be reversed. }

% In order to address \emph{this problem}, a new method has recently been
% proposed in \citet{held2020}. The approach combines the analysis of
% credibility \citep{matthews:2001,matthews:2001b} with a prior-data
% conflict assessment \citep{box:1980}. Replication success is declared if
% the replication study is in conflict with a sceptical prior that would
% make the original study non-significant. \emph{This approach penalizes
%   small relative effect sizes as we will see in more detail in the following.}

% To introduce some notation, let $z_o = \hat \theta_o/\sigma_o$ and
% $z_r = \hat \theta_r/\sigma_r$ denote the $z$-statistic of the
% original and replication study, respectively. Here $\hat \theta_o$ and
% $\hat \theta_r$ are the corresponding effect estimates (assumed to be
% normally distributed) of the unknown
% effect $\theta$ with standard errors $\sigma_o$ and $\sigma_r$, respectively.
% The corresponding one-sided
% $p$-values are denoted by $p_o=1-\Phi(z_o)$ and $p_r=1-\Phi(z_r)$,
% respectively, where $\Phi(\cdot)$ denotes the standard normal cumulative
% distribution function.
% Let $c = \sigma_o^2/\sigma_r^2$ denote the variance
% ratio of the squared standard errors of the original and replication
% effect estimates. The squared standard errors are usually inversely
% proportional to the sample size of each study, \ie
% $\sigma_o^2 = \kappa^2/n_o$ and $\sigma_r^2 = \kappa^2/n_r$ for some
% unit variance $\kappa^2$.  The variance ratio $c$ can then be
% identified as the relative sample size $c=n_r/n_o$. The relative effect size
% \begin{equation}\label{eq:d}
%   d = \frac{\hat \theta_r}{\hat \theta_o} = \frac{1}{\sqrt{c}} \frac{z_r}{z_o}
% \end{equation}
% quantifies the size of the replication effect estimate $\hat \theta_r$ relative
% to the original effect estimate $\hat \theta_o$. The corresponding shrinkage
% of the replication effect estimate will be denoted as $s=1-d$.

% Suppose the original study achieved statistical significance at one-sided
% level $\alpha$, so $p_o \leq \alpha$.  The standard approach to assess
% replication success is based on significance of the replication effect
% estimate at the same level $\alpha$, \ie the replication is considered successful if also $p_r \leq \alpha$.
% This approach is known in drug development as the two-trials rule \citep{senn:2007}, usually
% conducted at $\alpha=0.025$.
% Let $z_\alpha = \Phi^{-1}(1-{\alpha})>0$ denote the $z$-value corresponding
% to the level $\alpha$, then significance of the replication
% \emph{study is achieved if $z_r \geq z_\alpha$, which}
% is equivalent to the \emph{condition}
% \begin{equation}\label{eq:dSig}
%    d \geq \frac{z_\alpha}{z_o  \, \sqrt{c}}
% \end{equation}
% on the relative effect size \eqref{eq:d}.
% The right hand-side goes to zero for increasing $c$, so if the
% relative sample size $c$ is large enough, significance of the
% replication study can be achieved with any arbitrarily small (but
% positive) relative effect size $d$. \emph{However, declaring replication
% success when there is substantial shrinkage is contrary to common
% sense, as the replication effect estimate may not reflect an effect
% size of the same practical relevance as the original one, despite its
% statistical significance.}

% \emph{In this paper we first review the \citet{held2020} approach for
%   the assessment of replication success, followed by showing how it
%   relates to the relative effect size (Section \ref{sec:res}). This perspective
%   is used in Section \ref{sec:goldenthresh} and \ref{sec:recalib}
%   to propose  a recalibration of the
%   method, the \textit{golden level}, which leads to a more appropriate
%   criterion for replication success compared to the two-trials rule
%   (Section \ref{sec:2TR}). In Section \ref{sec:ER} we study
%   power and Type-I error rates of the proposed method and compare it to the
%   two-trials rule. The recalibrated method ensures that conditional power can
%   take any desired value if the original study has been significant and the replication
%   sample size is large enough
% (Section \ref{sec:powerrep}), controls the overall Type-I error if the replication
% sample size is not smaller than the original one (Section \ref{sec:T1E}), and
% offers uniform
% gains in project power compared to the two-trials rule (Section \ref{sec:PP}). Section \ref{sec:application}
% describes an application to data from four replication projects and Section
% \ref{sec:discussion} closes with some discussion.}

% \clearpage
% \section{Replication success}\label{sec:RS}

% \begin{center}
% \begin{figure}[!h]
% \begin{center}
% \begin{knitrout}
% \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

% {\centering \includegraphics[width=\maxwidth]{figure/fig1-1}

% }

% \end{knitrout}
% \caption{Example of the assessment of replication success. The original study from \citet{Pyc2010} has effect estimate  $\hat \theta_o=
%   0.4$ on Fisher's $z$ scale
%   (95\% CI from $0.05$ to $0.74$) and one-sided $p$-value
%   $p_o = 0.011$.
%   The left part of the figure illustrates the \emph{reverse-Bayes derivation of the sufficiently sceptical prior
%     based on}
%   the original study result and the posterior with lower credible limit fixed at zero. \emph{The comparison} of the sufficiently sceptical prior with the replication study result ($\hat \theta_r = 0.15$, 95\% CI from 0.04 to 0.26, $p_r = 0.004$)
%   in the right part of the figure \emph{is used to assess potential prior-data conflict.}
%   \label{fig:fig1}}
% \end{center}
% \end{figure}
% \end{center}

% Hereinafter we focus on the one-sided assessment of replication
% success to ensure that replication success can only occur if the
% original and replication effect estimates go in the same
% direction. Figure \ref{fig:fig1} illustrates the \citet{held2020} approach
% based on a replication study
% from the \textit{Social Sciences Replication Project} \citep{Camerer2018}:
% the significant original finding by \citet{Pyc2010} at
% one-sided level $\alpha=0.025$ is
% challenged with a sceptical prior, sufficiently concentrated around
% zero to make the original study result no longer convincing \citep{matthews:2001,matthews:2001b}.
% Replication success is then defined as conflict between the
% sceptical prior and the result from the replication study in order to
% \emph{disprove} the sceptic. Conflict is quantified by a prior-predictive
% tail probability $p_{\mbox{\scriptsize Box}}$ \citep{box:1980} where a small value
% $p_{\mbox{\scriptsize Box}} \leq \alpha$ defines replication success.
% In Figure \ref{fig:fig1}  the original finding is only borderline significant, so the
% sufficiently sceptical prior is fairly wide.  Furthermore, there is
% substantial shrinkage
% ($d = 0.15/0.4 =
% 0.38$)
% of the replication effect estimate
% and therefore hardly any conflict with the sufficiently
% sceptical prior (one-sided $p_{\mbox{\scriptsize Box}}=0.31$).
% We are thus not able to declare replication success at
% level $2.5$\%.

% \emph{The actual value of $p_{\mbox{\scriptsize
%       Box}}$ is difficult to interpret as it depends on the level $\alpha$ and does not even exist if
%   the original $p$-value $p_o$ exceeds $\alpha$. However, }
% \citet{held2020} showed
% that if both $\sign(z_o)=\sign(z_r)$ and
% \begin{equation}\label{eq:extrinsic.p}
% \left({z_o^2}/{z_{{\alpha_S}}^2}-1 \right) \left({z_r^2}/{z_{{\alpha_S}}^2}
% - 1 \right) \geq c
% \end{equation}
% hold, replication success at level ${\alpha_S}$ is achieved, where $z_{{\alpha_S}} = \Phi^{-1}(1-{\alpha_S})$.
% The requirement \eqref{eq:extrinsic.p} can be assessed for \emph{any} value of
% \emph{${{\alpha_S}} > \max\{p_o, p_r\}$} and of particular interest is the
% smallest possible value of $\alpha_S$ where \eqref{eq:extrinsic.p} holds,
% the so-called {\it sceptical $p$-value} $p_S$.
% We are thus interested in the value $z_S^2$ that fulfills
% \begin{equation}\label{eq:extrinsic.p2}
% \left({z_o^2}/{z_{S}^2}-1 \right) \left({z_r^2}/{z_{S}^2}
% - 1 \right) = c.
% \end{equation}
% There is a unique solution of \eqref{eq:extrinsic.p2} which defines
% the one-sided {sceptical $p$-value} $p_S=
% 1-\Phi\left({z_S}\right)$ where $z_{S} \coloneqq +
% \sqrt{z_{S}^2}$, provided
% $\sign(z_o)=\sign(z_r)$ holds.  Replication success at level
% $\alpha_S$ is then achieved if $p_S \leq
% \alpha_S$.  In the introductory example based on the original study by
% \citet{Pyc2010}, the sceptical
% $p$-value turns out to be $p_S=0.11$.

% The sceptical
% $p$-value has a number of interesting properties, see \citet[Section
% 3.1]{held2020} for details. In particular, $p_S > \max\{p_o, p_r\}$
% always holds with $p_S \downarrow \max\{p_o, p_r\}$ for $c \downarrow
% 0$. Furthermore, if the $p$-values $p_o$ and
% $p_r$ are fixed, the sceptical $p$-value
% $p_S$ increases with \emph{decreasing} relative effect size
% $d$.  The first property ensures that both the original and the
% replication study have to be sufficiently convincing on their own to
% \emph{achieve} replication success.  The second property guarantees that
% shrinkage of the replication effect estimate is penalized.

% The level for replication success $\alpha_S$ has to be distinguished
% from the significance level $\alpha$ associated with the ordinary
% $p$-value.  \citet{held2020} has used the {\it nominal level} for
% replication success ($\alpha_S=\alpha$) for convenience, but in the
% following we will propose a recalibration of the procedure along with
% a new value for $\alpha_S$, the {\it golden level} (Section~\ref{sec:goldenthresh}).
% \emph{The derivation is based on a property of the required
% relative effect size for replication success, if the relative sample size is very large
% (Section~\ref{sec:res}).}
% In a nutshell, the golden level
% ensures that for original studies which were only borderline
% significant \emph{($p_o=\alpha$)}, replication success is only possible if the replication
% effect estimate is larger than the original one \emph{($d > 1$)}.\\

% \subsection{Relative effect size}\label{sec:res}
% Without loss of generality we \emph{now} assume that $\hat \theta_o > 0$ and that
% $p_o < {\alpha_S}$ has been observed in the original study, otherwise
% it would be impossible to achieve replication success at level
% $\alpha_S$ because $p_S$ is always larger than $p_o$.
% The condition \eqref{eq:extrinsic.p} for replication
% success can then be re-written as
% \begin{equation}\label{eq:eq.z}
%   z_r \geq   z_{{\alpha_S}} \sqrt{1+c/(K-1)} \eqqcolon  \zrmin,
% \end{equation}
% where $K=z_o^2/z_{{\alpha_S}}^2>1$. The right hand-side of
% \eqref{eq:eq.z} is the minimum replication $z$-value $\zrmin$ required
% to achieve replication success.  Note that $\zrmin$ increases with increasing $c$, so increasing the replication
% sample size leads to a more stringent success requirement \emph{for $z_r$ and the corresponding} replication
% $p$-value $p_r$.

% Equation \eqref{eq:eq.z} can be further transformed to a
% condition on the relative effect size \eqref{eq:d}:
% \begin{equation}\label{eq:res}
%   d \geq  \frac{\sqrt{1+c/(K-1)}}{ \sqrt{c K}} \eqqcolon \dmin .
% \end{equation}
% To achieve replication success, the relative effect size
% must be
% at least as large as the right hand-side of \eqref{eq:res}, the
%   minimum relative effect size $\dmin$, a function of $K$ and the
% relative sample size $c$.
% If the relative sample
% size becomes very large, \ie $c \rightarrow \infty$, we have
% $\dmin \downarrow \dinfty$ where
% \begin{equation}\label{eq:smallestpossible}
%   \dinfty = 1/\sqrt{K (K-1)}
% \end{equation}
% \emph{is the {\it limiting relative effect size}.}
% This shows that the minimum relative effect size $\dmin$ in \eqref{eq:res}
% does not go to zero for increasing $c$, so replication success cannot
% be achieved if the relative effect size $d$ is smaller or equal to
% $\dinfty$, no matter how large the replication study is.
% \emph{In contrast}, the corresponding criterion \eqref{eq:dSig} \emph{of the
% two-trials rule}
% can be achieved for any positive relative effect size, regardless of
% how small, provided the replication sample size is sufficiently large.

% \subsection{The golden level}\label{sec:goldenthresh}
% Significance of both the original and the replication study at level
% $\alpha$ is a necessary but not sufficient requirement for replication
% success at the nominal level ($\alpha_S = \alpha$).  The nominal level
% may therefore be too stringent.  It is more reasonable to calibrate
% the procedure in such a way that to establish replication success,
% original and replication study do not both necessarily need to be
% significant at level $\alpha$, provided that the replication effect
% estimate does not shrink compared to the original one.  We therefore
% choose a level $\alpha_S$ such that a borderline
% significant original study ($p_o = \alpha$) cannot lead to replication
% success if there is shrinkage $s > 0$ of the replication effect
% estimate.  Mathematically, this translates to setting $\dinfty = 1$
% and $K = z_\alpha^2/z_{\alpha_S}^2$ in~\eqref{eq:smallestpossible} and
% leads to the quadratic equation $K(K-1) = 1$ with solution
% $K = \varphi$ where
% $\varphi = (\sqrt{5}+1)/2 \approx 1.62$ is
% known as the golden ratio. Solving for $z_{\alpha_S}$ gives
% \emph{$z_{\alpha_S} = z_\alpha / \sqrt{\varphi}$
% and the corresponding  \textit{golden level}
% \begin{eqnarray}\label{eq:golden}
% \alpha_S & = & 1 -  \Phi(z_\alpha / \sqrt{\varphi} )
% \end{eqnarray}
% for replication success. }
% This is our recommended default choice to assess replication success and we will study its properties in the following
%                    in more detail. For $z_\alpha = 1.96$ (one-sided $\alpha = 0.025$),
% the golden level is $\alpha_S =0.062$.
% In the introductory example
% shown in Figure \ref{fig:fig1},
% the sceptical $p$-value is $p_S=0.11 > 0.062$, so
% the replication  of the \citet{Pyc2010} study was not successful.

% \emph{The golden level \eqref{eq:golden} is derived from~\eqref{eq:smallestpossible} with $\dinfty = 1$.
% However, we may also use a different value for the
% limiting relative effect size $\dinfty$, say $\dinfty=0.8$.
% Then replication success is only possible for a borderline significant result ($p_o = \alpha$)
% if there is less than $1-\dinfty$ (20\% for $\dinfty=0.8$) shrinkage of the replication effect estimate.
% This approach is equivalent to a limiting relative effect size of 1 if the original $p$-value
% $p_o$ is equal to a different level $\alpha'$, which can be derived as follows:
% First, solving \eqref{eq:smallestpossible} for $\dinfty > 0$ gives
% $K = z_\alpha^2/z_{\alpha_S}^2 = 1/2+\sqrt{1/4 + 1/\dinfty^2}$. The new level
% $\alpha'$ fulfills  $\varphi=z_{\alpha'}^2/z_{\alpha_S}^2$, so $z_\alpha^2/K=z_{\alpha'}^2/\varphi$
% and therefore
% \begin{equation}\label{eq:alphaPrime}
% \alpha' = 1-\Phi\left(z_{\alpha} \, \sqrt{\varphi / K} \right).
% \end{equation}
% For example, for $\alpha = 0.025$ and $\dinfty = 0.8$ we obtain $\alpha'=0.033$.
% }

% \subsection{Recalibration of the sceptical $p$-value}\label{sec:recalib}

% The condition $p_S \leq \alpha_S$ for replication success at the
% golden level is equivalent to $z_S \geq z_\alpha / \sqrt{\varphi}$,
% \ie $z_S \sqrt{\varphi} \geq z_\alpha$.  In practice it may be
% preferable to recalibrate the sceptical $p$-value
% $p_{S} = 1 - \Phi(z_S)$ to
% $ \tilde{p}_S = 1 - \Phi(z_S \sqrt{\varphi})$, which then needs to be
% compared to $\alpha$ (rather than $\alpha_S$) to assess replication
% success and can thus be interpreted on the same scale as an ordinary
% $p$-value.  For example, the recalibrated sceptical $p$-value for the
% replication of \citet{Pyc2010} turns out to be
% $\tilde{p}_S=0.061$ and does not lead to replication
% success at \emph{any level $\alpha<0.061$, including} the
% standard $0.025$ level.

% \subsection{Comparison with the two-trials rule}\label{sec:2TR}
% A useful benchmark for comparison is the two-trials rule in drug
% development \citep[Section 9.4]{kay:2015}, which requires ``at least
% two adequate and well-controlled studies, each convincing on its own,
% to establish effectiveness'' \citep[p.~3]{FDA1998}.  This is usually
% achieved by independently replicating the result of a first study in a
% second study, both significant at one-sided level
% $\alpha=0.025$.  {It is worth noting that in practice the two trials are often run in parallel} {\citep{senn:2007}}{, so do not exactly resemble the replication setting.}

% The main difference between the replication success and the two-trials
% rule approach concerns how shrinkage of the replication effect
% estimate is handled. Figure~\ref{fig:fig2} illustrates that
% shrinkage is penalized in the assessment of replication success, \ie
% the original $p$-value needs to be quite small to achieve replication
% success for a relative effect size $d<1$. In contrast, significance of
% the replication study can be achieved even if there is substantial
% shrinkage, provided the replication sample size is large enough.

% \begin{figure}[!h]
% \centering
% \begin{knitrout}
% \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

% {\centering \includegraphics[width=\maxwidth]{figure/fig2-1}

% }

% \end{knitrout}
% \caption{Comparison of replication success at the golden level ($p_S \leq \alpha_S =
% 0.062$) and the two-trials rule ($p_o \leq 0.025$ and $p_r \leq 0.025$).
% The \emph{dotted} areas indicate that success is
% impossible for original $p$-value $p_o$ and relative effect size $d$.
% In the white areas success is possible and
% depends on the relative sample size $c$ as indicated by the grey lines.
% \emph{The dashed black line in the left plot indicates the limiting relative
% effect size $\dinfty$.}
% }
% \label{fig:fig2}
% \end{figure}

% It is interesting to directly compare the two-trials rule and
% replication success at the golden level in terms of
% the required relative effect size $d$ to fulfill the criteria
% \eqref{eq:dSig} and \eqref{eq:res}, respectively, see Figure~\ref{fig:fig2}.  If the original
% $p$-value is not significant at level $\alpha$, only replication
% success can be achieved, but will require a replication effect
% estimate larger than the original one.
% \emph{
%   For example, four studies with one-sided
%   $p_o \in (0.025, 0.03)$ have been included in the
%   \textit{Reproducibility Project: Psychology} \citep{OSC2015} and one
%   of them achieves replication success (see Section \ref{sec:application} for
%   details).  By definition, such non-significant original findings can
%   never fulfill the two-trials rule.
% }

% If the original $p$-value is
% smaller than $\alpha$, then the situation depends on the relative
% sample size $c$.  For example, when the replication sample size is
% chosen to be the same as in the original study ($c=1$) and
% $\alpha=0.025$, original studies with a $p$-value larger than
% $0.006$ will require a smaller relative effect
% size $d$ with the two-trials rule, while $p$-values smaller than
% $0.006$ will require a smaller relative effect
% size $d$ with the replication success method.  This illustrates that
% the latter method is less stringent than the two-trials rule if the
% original study is already sufficiently convincing.

% \section{Power and Type-I Error Rate}\label{sec:ER}
% Although Bayesian methods do not rely on the frequentist paradigm of
% repeated testing, it is still useful to investigate their frequentist
% operating characteristics \citep{Dawid1982, Rubin1984, Grieve2016}
% and this also holds for the proposed reverse-Bayes assessment of
% replication success. We first condition on the results from the original
% study and compare the power to achieve replication success with the two-trials
% rule in Section \ref{sec:powerrep}.
% We then
% assume that none of the two studies have been conducted and
% investigate the \emph{overall} Type-I error rate (Section \ref{sec:T1E}) and the project power
% (Section \ref{sec:PP})
% \citep{Maca2002} \emph{over both studies in combination for fixed relative sample size $c$}.

% \subsection{Conditional power}\label{sec:powerrep}

% \begin{figure}[!ht]
% \begin{center}
% \begin{knitrout}
% \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

% {\centering \includegraphics[width=\maxwidth]{figure/fig3-1}

% }

% \end{knitrout}
% \end{center}
% \caption{\emph{Conditional} power as a function of the {one-sided}
%   $p$-value of the original study with \emph{relative} sample size
%   $c=1$ (left) and $c=5$ (right). Shown is
%   conditional power \emph{assuming the unknown parameter is equal to
%   the original effect estimate (solid)} \emph{and conditional
%     power based on 20\% shrinkage of the original
%     effect estimate (dashed)} for the two-trials rule (2TR) at level
%   $\alpha = 0.025$ and for replication success
%   % (RS)
%   at the
%   corresponding golden and nominal level. Power values of exactly zero
%   are omitted.}
% \label{fig:fig3}
% \end{figure}

% Figure~\ref{fig:fig3} compares the power for replication success
% \emph{\citep[see][Section 4 for details]{held2020}} at the golden and at
% the nominal level with the power of the two-trials rule for relative
% sample size $c = 1$ (left) and $c = 5$ (right)
% as a function of the one-sided $p$-value \emph{$p_o$} from the original
% study.  \emph{Shown is the conditional power assuming the unknown parameter
%   $\theta$ is equal to the original effect estimate $\hat \theta_o$. Then
% $\hat \theta_r \given \hat \theta_o \sim \Nor(\hat \theta_o, \kappa^2/n_r)$
% and it follows that
% $d \given \hat \theta_o \sim \Nor(1, 1/(c z_o^2))$.
% The conditional power for replication
% success can therefore be calculated as
% \begin{equation}\label{eq:condPower}
% \P(d \geq \dmin \given \hat \theta_o) = \Phi\left[\sqrt{c} z_o (1-\dmin)\right],
% \end{equation}
% where $\dmin$ is
% given in \eqref{eq:res}.  Predictive power, which is conditional
% power averaged over a $\Nor(\hat \theta_o, \sigma_o^2)$ distribution
% for the effect size $\theta$, could also be calculated, then
% \mbox{$d \given \hat \theta_o \sim \Nor(1, (1+1/c)/z_o^2)$}.
% Conditional and predictive power of the two-trials rule
% also depend on $z_o$, $c$ and $\alpha$ and are given in
% \citet{MicheloudHeld2021}.}

% The two-trials rule requires a significant original study and hence it is
% impossible to power a replication study when $p_o > 0.025$.
% \emph{The same applies for replication success at the nominal level,
% where the power is zero for any $p_o > 0.025$, regardless
% of the replication sample size.
% This is different for the golden level,
% where the conditional power of an original study with $0.025 < p_o
% < 0.062$ is low, but not zero.}
% \emph{However, if the original $p$-value $p_o$ is slightly smaller than
% $0.025$, the two-trials rule has a larger power, both for $c=1$ and $c=5$. }
% \emph{But if the original $p$-value is sufficiently
% small ($p_o < 0.006$ for $c=1$), the power for replication success at the golden level is larger
% than the power of the two-trials rule. }

% \emph{ Compared to
%   $c=1$, the conditional power for $c=5$
%   of both the two-trials rule
%   and the replication success approach at the golden level
%   increases if $p_o \leq \alpha$.     A remarkable feature of the
%   replication success approach at the golden level is that conditional
%   power can be pushed towards 100\% for large enough $c$ if
%   $p_o<\alpha$, but not otherwise. This can be seen from \eqref{eq:condPower} because
%   $\dmin < 1$ for $p_o<\alpha$ and large enough relative sample size $c$.
% On the other hand, for $p_o > \alpha$ conditional power for replication success
% will tend to 0\% for increasing $c$ because $\dmin>1$ for all $c$. Finally, for $p_o = \alpha$
% the limit is 50\%.
%   The same
%   property can be observed at the nominal level, however at the
%   smaller threshold $1-\Phi(z_\alpha \sqrt{\varphi})$ which is $0.006$ for  $\alpha=0.025$. Only if $p_o < 0.006$ will the conditional power
%   for replication success attain 100\% for $c \to \infty$. This further highlights the stringency
%   of the nominal level.  }

% \emph{The approach described so far takes the original study at
%   face-value since it assumes that $\hat \theta_o$ is equal to the unknown effect size $\theta$.
%   In practice, however,
%   there are often good reasons to believe that original effect estimates
%   have a tendency to be inflated (\eg due to publication bias). One
%   way to address this issue is to base power calculations on a
%   shrunken version of the original effect
%   estimate, where the amount of shrinkage is guided by domain knowledge and
%   a risk of bias assessment of the original study.
%   For illustration, Figure~\ref{fig:fig3} also shows conditional power based
%   on 20\% shrinkage of the original effect estimate
%   which reduces the conditional power for all methods, especially for
%   a relative sample size $c = 1$.  Conditional power for
%   replication success at the golden level can now be pushed towards 100\%
%   only for $p_o < 0.018$, which can be derived by
%   solving \eqref{eq:alphaPrime} for $\alpha$ with
%   $\alpha'=0.025$ and $\dinfty=0.8$.  To be able to
%   push conditional power based on 20\% shrinkage towards 100\% for all
%   $p_o < 0.025$,
%   equation \eqref{eq:alphaPrime} would have to be used directly to relax the level
%   from $\alpha=0.025$ to $\alpha'=0.033$.
%   }

% \subsection{Overall Type-I error rate}\label{sec:T1E}
% \emph{The two studies are assumed to be independent with Type-I error
% rate fixed at $\alpha$ for each of them,} so the Type-I error rate of the
% two-trials rule \emph{over the entire project} is simply $\alpha^2$ for
% any value of the relative effect size $c$.
% In contrast, the Type-I
% error rate of the proposed replication success assessment {depends on}
% \emph{the relative sample size} $c$.

% For $c=1$, \citet[Section 3]{held2020} showed that $z_S^2$ in \eqref{eq:extrinsic.p2}
% simplifies to half the harmonic mean of the squared test statistics
% $z_o^2$ and $z_r^2$.  The
% connection $z_S^2 = z_H^2/4$ to the harmonic mean $\chi^2$-test
% statistic $z_H^2$ \citep{held2020b}, which has a
% $\chi^2(1)$-distribution under the null hypothesis, makes it
% straightforward to compute the Type-I error rate  at level
% $\alpha_S$ for $c = 1$ as
% \begin{eqnarray}\label{eq:T1E}
% \mbox{T1E} =   \left\{1-\Phi\left[2 \, \Phi^{-1}\left(1-\alpha_S \right)
%  \right]\right\}/2.
% \end{eqnarray}
% For the golden level $\alpha_S =0.062$ at
% $\alpha = 0.025$, the Type-I error rate \eqref{eq:T1E}
% is $0.0515$\%, slightly less than the Type-I
% error rate $\alpha^2=0.0625$\% of the
% two-trials rule.  For comparison, the Type-I error rate at the nominal
% level $\alpha_S=0.025$ is
% $0.0022$\%, \emph{much} smaller than
% 0.0625\%.

% For $c \neq 1$,  the Type-I error rate can be calculated through numerical
% integration:
% \begin{equation}\label{eq:T1E0}
%   \mbox{T1E} = \int_{z_{\alpha_S}}^{\infty}
% \P(z_r \geq \zrmin \given z_o, c, \alpha_S) \,
%   \phi(z_o) \, dz_o,
% \end{equation}
% where $\phi(\cdot)$ denotes the standard normal density function.
% The first term in the integral of \eqref{eq:T1E0} is the probability of
% replication success at level $\alpha_S$ conditional on a fixed original test
% statistic $z_o$ and a relative sample size $c$.
% Now $z_r \sim \Nor(0,1)$ under the null hypothesis, so
% this term simplifies to
%   $\P(z_r \geq \zrmin \given z_o, c, \alpha_S) = 1 - \Phi(\zrmin)$
% where $\zrmin$ in \eqref{eq:eq.z} depends on $z_o$, $c$, and $\alpha_S$.

% \begin{figure}[!h]
% \centering

% \begin{knitrout}
% \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

% {\centering \includegraphics[width=\maxwidth]{figure/fig4-1}

% }

% \end{knitrout}

% \caption{\emph{Overall} Type-I error rate (left) and project power (right) for fixed relative sample size $c$.
%   Results are given for replication success
%   % (RS)
%   at the nominal and golden
%   level and compared with the two-trials rule (2TR) at $\alpha = 0.025$.
%   \emph{The dashed darkgrey line is the project power at the golden level based on significant original studies ($p_o \leq 0.025$).}
%   The power of the original study is 90\%.}
% \label{fig:fig4}
% \end{figure}

% The left plot in Figure \ref{fig:fig4} displays the Type-I error
% rate for $\alpha=0.025$ as a function of the relative sample
% size $c$.  It can be seen that the Type-I error \emph{of the replication success approach} decreases with
% increasing relative sample size $c$. \emph{This also follows from \eqref{eq:T1E0}
%   where
% $\P(z_r \geq \zrmin \given z_o, c, \alpha_S) = 1 - \Phi(\zrmin)$
% decreases with increasing $c$, because $\zrmin$ increases with increasing $c$,
% see equation \eqref{eq:eq.z}. }

% The Type-I error rate of the nominal level is always below the target
% $0.0625$\%.  Although the Type-I error will
% eventually attain $\alpha^2$ in the limit $c \downarrow 0$
% \citep[Section 3.4]{held2020}, the nominal level seems to be too
% stringent for realistic values of $c$.  The Type-I error rate of the
% golden level is smaller than $0.0625$\% for
% $c >0.85$.  Appropriate Type-I error control is
% thus ensured even for replication studies where the sample size is
% slightly smaller than in the original study.

% \begin{figure}[!ht]
% \begin{center}

% \begin{knitrout}
% \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

% {\centering \includegraphics[width=\maxwidth]{figure/fig5-1}

% }

% \end{knitrout}

% \end{center}
% \caption{\emph{Overall Type-I error rate if the
%   replication sample size equal to the original study
%   ($c=1$).  The two-trials rule (2TR)
%   is compared to replication success
%   % (RS)
%   at the
%   golden and nominal level for different values of $\alpha$.}}
% \label{fig:fig5}
% \end{figure}

% \emph{Figure \ref{fig:fig5} compares for $c=1$ the Type-I error rate \eqref{eq:T1E}
%   of replication success at the golden and at the nominal level
%     with the
%   two-trials rule for different values of $\alpha$. The
%   Type-I error rate of the two-trials rule is $\alpha^2$ and the
%   replication success approach at the nominal level always has a much
%   smaller Type-I error rate than $\alpha^2$.  At the golden level
%   the Type-I error rate of the
%   replication success approach is much closer to $\alpha^2$, still slightly smaller
%   if $\alpha < 0.058$. For $\alpha = 0.058$ the
%   Type-I error rate is equal to the
%   Type-I error rate $0.058^2 = 0.34\%$ of the
%   two-trials rule and for $\alpha > 0.058$ the
%   Type-I error rate is slightly larger than $\alpha^2$.
%   The
%   Type-I error rate for replication success decreases with
%   increasing $c$, so as long as the replication sample size is not
%   smaller than the original sample size, Type-I error control at
%   $\alpha^2$ is guaranteed at the golden level for any one-sided level
%   $\alpha < 0.058$.}

% \subsection{Project power}\label{sec:PP}
% Under the alternative we have
% $z_o\sim \Nor(\mu, 1)$ with $\mu = z_\alpha + z_\beta$
% where $\alpha$ is
% the assumed significance level and
% \emph{
% $1-\beta = \Phi(\mu - z_\alpha)$
% is the power to detect the
% assumed effect $\theta = \mu \sigma_o$ in the original study \citep[Section 3.3]{mat2006}.
% In the following
% $\alpha=0.025$ and $\beta=0.1$ are
% used. The power of a significant
% replication study with sample size $n_r = c n_o$ is
% \begin{equation*}
% \Phi(\theta/\sigma_r - z_\alpha) =
% \Phi(\sqrt{c} \mu - z_\alpha),
% \end{equation*}
% }
% so
% depends on $\mu$ and the relative sample size $c$. The project power of the
% two-trials rule is therefore $(1-\beta) \, \Phi(\sqrt{c} \mu - z_\alpha)$ \emph{and increases with increasing $c$.}

% The project power for replication success is computed as
% \begin{equation*}
%   \mbox{PP} = \int_{z_{\alpha_S}}^{\infty}
% \P(z_r \geq \zrmin \given z_o, c, \alpha_S) \,
%   \phi(z_o - \mu) \, dz_o
% \end{equation*}
% and shown in the right plot of Figure \ref{fig:fig4} as a function
% of $c$.  For the golden level, the project power quickly increases to
% values above 90\%, whereas the nominal level only reaches around 80\%
% project power. The project power based on the two-trials rule is shown
% for comparison, which is always smaller than for the golden level
% and converges to 90\% for large $c$.

% \emph{The advantage in power stems partly from replication success still
%   being possible when the original $p$-value is larger than 0.025, but
%   smaller than 0.062. If we assume that a replication study is
%   only conducted if the original study is significant (with
%   $p_o \leq 0.025$), then the project power based on the
%   golden level (the dashed line in Figure \ref{fig:fig4}) is
%   slightly smaller and for $c>1$ barely different than for the
%   two-trials rule.  More substantial gains are still visible for
%   $c < 1$.
%   However, the restriction to original studies with
%   $p_o \leq 0.025$ may not reflect current practice in
%   large-scale replication projects. For example, 5 out of
%   143 replication studies considered in Section
%   \ref{sec:application} do have original $p$-values between
%   0.025 and 0.062.  }

% \section{Application}\label{sec:application}
% In this section, we illustrate the proposed methodology using
% data from four replication projects.
% All four projects reported effect estimates that were transformed to
% correlation coefficients ($r$). This scale allows for easy comparison of
% effect estimates from studies that investigate different phenomena
% \emph{and} is bounded to the interval between minus one and one.
% Moreover, the Fisher $z$-transformation $\hat{\theta} = \text{tanh}^{-1}(r)$
% can be applied to the correlation coefficients, resulting in the transformed
% estimates being asymptotically normal with variance which is only a function
% of the study sample size $n$, \ie $\Var(\hat{\theta}) = 1/(n - 3)$
% \citep{Fisher1921}.

% The first data set comprises the results from the \textit{Reproducibility Project:
% Psychology} \citep{OSC2015}, whose aim was to replicate 100 studies, all of which
% were published in three major Psychology journals in 2008. For our purpose only
% the 73 study pairs from the ``meta-analytic'' subset are considered, since only
% for these studies the standard error of the Fisher $z$-transformed effect
% estimates can be computed \citep{Johnson2016}.
% The second data set comes from the \textit{Experimental Economics Replication
% Project} \citep{Camerer2016} which attempted to replicate 18 experimental
% economics studies published in two high impact economics journals between 2011
% and 2015.
% The third data stem from the \textit{Social Sciences Replication Project}
% \citep{Camerer2018} where 21 replications of studies on the social sciences
% were carried out, all of which were originally published in the journals
% \textit{Nature} and \textit{Science} between 2010 and 2015.
% The last data set originates from the \textit{Experimental Philosophy Replicability
% Project} \citep{Cova2018} which involved 40 replications of studies from the
% emerging field of experimental philosophy. Since only for 31 studies effective
% sample size for original and replication study were available simultaneously,
% only these pairs were included. For more information on the data sets see
% also \citet{Pawel2020}.

% Table~\ref{tab:marginalres} presents overall results for each of the
% replication projects. While the median relative effect size is below one for
% all of the four projects, there are still large differences. For
% example, the median relative effect size is only
% $0.29$
% in the Psychology project, whereas it is
% $0.86$ in the Philosophy project. The degree of shrinkage is also
% reflected in the success rates (according to the two-trials rule and the replication
% success approach \emph{at the golden level}), which are around 30\% for the former and more than 70\%
% for the latter. The proportion of successful replications is similar for the
% two-trials rule and the replication success approach. \emph{In the Experimental
% Economics project the methods perfectly agree, while in the other three projects
% the methods disagree for a few studies.}

% \begin{table}[!ht]
% \caption{Results for each replication project:
% Relative effect size $d$ (median with 25\% and 75\% quantiles on Fisher's $z$ scale), proportion of
% successful replications with the two-trials rule (2TR) and the replication success (RS)
% approach (at the golden level), and number of studies where the methods disagree.}
% % \resizebox{\textwidth}{!} {
% % latex table generated in R 4.0.5 by xtable 1.8-4 package
% % Fri Nov 25 10:07:05 2022
% \begin{tabular}{lcccc}
%   \toprule
% Project & relative effect size $d$ & 2TR (\%) & RS (\%) & discrepant \\
%   \midrule
% Psychology & 0.29 [0.03, 0.77] & 28.8 & 30.1 & 3/73 \\
%   Experimental Economics & 0.67 [0.35, 0.92] & 55.6 & 55.6 & 0/18 \\
%   Social Sciences & 0.52 [0.13, 0.65] & 61.9 & 52.4 & 2/21 \\
%   Experimental Philosophy & 0.86 [0.47, 1.12] & 74.2 & 71.0 & 1/31 \\
%    \bottomrule
% \end{tabular}

% % }
% \label{tab:marginalres}
% \end{table}

% \begin{figure}[!h]
% \centering
% \begin{knitrout}
% \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

% {\centering \includegraphics[width=\maxwidth]{figure/fig6-1}

% }

% \end{knitrout}
% \caption{Relative effect size $d$ versus original $p$-value $p_o$.
% Black indicates that replication success was achieved at the golden level
% while grey indicates that it was not. The diamonds mark studies where
% the replication success approach (at the golden level) and the two-trials rule disagree. The dashed
% black line indicates the \emph{limiting relative effect size}
% at the golden level with $\alpha = 0.025$.
% }
% \label{fig:fig6}
% \end{figure}

% Figure~\ref{fig:fig6} displays the relative effect size $d$ versus the
% original $p$-value $p_o$ for each study pair and stratified by project.
% Note that one study pair from the Philosophy project is not shown due to
% extremely small original $p$-value and another study pair from the Psychology project
% is not shown due to a very large relative effect size.
% We can see that for most of the study pairs, the replication success approach
% and the two-trials rule lead to the same conclusion, only six replications
% show conflicting results.
% They are highlighted with diamonds in Figure~\ref{fig:fig6} and their
% characteristics are summarised in Table~\ref{tbl:discrep}.
% \begin{table}[!ht]
% \caption{Characteristics of studies for which the replication success approach
% (at the golden level)
% and the two-trials rule disagree (at one-sided $\alpha = 0.025$). Shown are relative
% sample size $c$, relative effect size $d$, original, replication and recalibrated sceptical
% $p$-value $p_o$, $p_r$ and $\tilde{p}_S$.}
% \label{tbl:discrep}
% % \resizebox{\textwidth}{!} {
% % latex table generated in R 4.0.5 by xtable 1.8-4 package
% % Fri Nov 25 10:07:05 2022
% \begin{tabular}{lllllll}
%   \toprule
% Study & Project & $c$ & $d$ & $p_o$ & $p_r$ & $\tilde{p}_S$ \\
%   \midrule
% \citet{Schmidt2008} & Psychology & 2.58 & 1.28 & \textbf{0.028} & \textcolor{black}{< 0.0001} & \textcolor{black}{0.024} \\
%   \citet{Oberauer2008} & Psychology & 0.60 & 0.67 & \textcolor{black}{0.0003} & \textbf{0.035} & \textcolor{black}{0.017} \\
%   \citet{Payne2008} & Psychology & 2.65 & 0.41 & \textcolor{black}{0.001} & \textcolor{black}{0.023} & \textbf{0.031} \\
%   \citet{Balafoutas2012} & Social Sciences & 3.48 & 0.52 & \textcolor{black}{0.009} & \textcolor{black}{0.011} & \textbf{0.04} \\
%   \citet{Pyc2010} & Social Sciences & 9.18 & 0.38 & \textcolor{black}{0.011} & \textcolor{black}{0.004} & \textbf{0.061} \\
%   \citet{Nichols2006} & Experimental Philosophy & 9.40 & 0.49 & \textcolor{black}{0.015} & \textcolor{black}{0.0006} & \textbf{0.049} \\
%    \bottomrule
% \end{tabular}

% % }
% \end{table}
% Two studies from the Psychology project show replication success
% but fail the two-trials rule.
% These studies show $p$-values that are slightly above the significance
% threshold in either original or replication study, but do not exhibit much
% shrinkage; in
% the replication of \citet{Oberauer2008}, the replication $p$-value was
% $p_r = 0.035$,
% a little too large to pass the two-trials rule. However, as the replication
% effect estimate shrunk only about
% $30$\%
% compared to the original one, replication success is still achieved. Conversely,
% the original $p$-value
% $p_o = 0.028$
% in \citet{Schmidt2008} was just above the significance level, yet the
% replication led to a highly significant result
% $p_r < 0.0001$
% with the effect estimate being even
% $30$\%
% larger than the original counterpart, which therefore also resulted in
% replication success.

% The remaining conflicting studies do not show replication
% success despite passing the two-trials rule.
% In all cases, there is substantial shrinkage of the replication effect estimate
% compared to the original one. For instance, in the replication study of
% \citet{Pyc2010}, the estimate shrunk by
% $62$\%
% and the replication $p$-value was only significant because the
% sample size was increased by a factor of
% $c = 9.2$.

% \emph{This analysis was based on the default choice $\dinfty=1$
%   at $\alpha=0.025$ for the golden level as described in Section
%   \ref{sec:goldenthresh}. We may also choose a different value for the
%   limiting relative effect size $\dinfty$ at $\alpha=0.025$ which then
%   corresponds to $\dinfty=1$ at a different level $\alpha'$ as given in \eqref{eq:alphaPrime}.  Figure
%   \ref{fig:fig7} compares the proportion of successful
%   replications with the replication success approach for
%   $\dinfty \in (0.5, 1.1)$ with the two-trials rule at the
%   corresponding levels
%   $\alpha' \in (0.06,
%   0.022)$ for all four replication projects.
%   We can see that the two proportions agree fairly well for all values
%   of $\alpha'$ considered. The number of discrepant studies in each project varies
%   between 0 and 3.   Only in the Psychology project there are
%   some studies which are successful with the replication success approach but
%   not the two-trials rule
%   and some studies successful with the two-trials rule but not the replication
%   success approach.  The proportion of studies where both
%   methods are successful (also shown in Figure \ref{fig:fig7}) is then smaller than the proportion of
%   successful replications with either one of the two methods.  The
%   three discrepant studies  from the Psychology project listed in the top three rows of
%   Table~\ref{tbl:discrep} are an example of this particular feature.
% }

% \begin{figure}[!htb]
% \begin{knitrout}
% \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

% {\centering \includegraphics[width=\maxwidth]{figure/fig7-1}

% }

% \end{knitrout}
% \caption{Proportion of successful replications as a function of the limiting
% relative effect size $\dinfty$ at $\alpha = 0.025$. The upper axis gives the
% equivalent level $\alpha'$ where the corresponding limiting relative effect size is 1.
% The replication success (RS) approach is compared with the two-trials rule (2TR).}
% \label{fig:fig7}
% \end{figure}

% \section{Discussion}\label{sec:discussion}
% In this paper, we have expanded on the replication success approach introduced
% in \citet{held2020} and demonstrated its advantages over alternative methods
% such as the two-trials rule. In particular, the method provides an attractive
% compromise between hypothesis testing and estimation,
% as it penalizes
% shrinkage of the replication effect estimate compared to the original one,
% while ensuring that both are statistically significant to some extent.
% %% For instance, the method will indicate only a low degree of replication success when
% %% the replication study shows a much smaller but statistically significant
% %% effect estimate, whereas it can still indicate some degree of
% %% success when either original or replication $p$-value are slightly above the
% %% significance level. %% , provided their effect estimates are compatible.
% %% %% \todo{SP: maybe ``provided the replication estimate does not shrink?''}

% We further refined the method by proposing the golden level, a
% new threshold for replication success. It guarantees that borderline
% significant original studies can only be replicated successfully if
% the replication effect estimate is larger than the original
% one. Compared to the two-trials rule, the golden level offers uniform
% gains in project power and controls the Type-I error rate \emph{at any
%   one-sided level $\alpha < 0.058$} if the
% replication sample size is not \emph{smaller than the original one}.
% Empirical evaluation of data from four replication projects highlights
% that in most cases the methods are in agreement, however, for the
% study pairs where the approaches disagree, the replication success
% approach seems to lead to more sensible conclusions.
% \emph{The good performance has been recently confirmed by a comparison of different
%   replication success metrics through a simulation study in the
%   presence of publication bias \citep{Muradchanian2021}.}

% Despite a lack of agreement as to which statistical method should be used to
% evaluate replication studies, conclusions based on different methods usually
% agree. Nevertheless, in some cases, classical methods such as the two-trials
% rule may produce anomalies. We argue that
% the replication success approach improves upon existing methods leading to
% more appropriate inferences and decisions that better
% reflect the available evidence.
% \emph{However, in extreme cases the performance of the sceptical $p$-value may be
%   considered as strange or even counterintuitive. Specifically, if the
%   original study was only borderline significant, a highly significant
%   replication study can only lead to success if the replication effect
%   estimate is larger than the original one. To understand this
%   behaviour it is important to realize that the proposed approach does
%   not synthesize the evidence from the two studies (like a standard
%   meta-analysis). The sceptical $p$-value is designed to confirm
%   claims of new discoveries through replication, but will remain
%   ``stubborn'' \citep{LyWagenmakers2020} if the original study was not
%   particularly convincing, even if the replication study
%   provides overwhelming evidence for an effect.  It will lead to a
%   different result if the order of studies was reversed, as long as
%   original and replication study do not have the same sample size
%   ($c \neq 1$).  The related harmonic mean $\chi^2$-test
%   \citep{held2020b} for evidence synthesis of two or more studies
%  also requires  each study to be convincing on its own to a certain degree, but
%   treats them as exchangeable. }

% With this paper we further advanced the reverse-Bayes methodology for
% the analysis and design of replication studies, yet certain
% limitations and opportunities for future research remain: First,
% assuming normality of the effect estimates may be questionable,
% especially for small sample sizes, and more robust distributional
% assumptions could be considered.  Second, in some types of analyses
% (\eg regression or ANOVA) the effect estimate is a vector and the
% approach would need suitable adaptations.
% Third, there is a recent trend to not only conduct one but
% several replications for one original study \citep[\eg][]{Klein2014, Ebersole2016, Klein2018}.
% Also for this situation,
% the method would need to be adapted, \eg the replication estimates
% could be first synthesized and an analysis of replication
% success could be performed subsequently.

% \emph{Throughout the paper we have assumed that the relative sample size is
% fixed in advance.  In practice the sample size of the replication
% study is often chosen based on the result of the original study
% \citep{Anderson2017}.  Power calculations as shown in Figure
% \ref{fig:fig3} can then be inverted to determine the appropriate
% sample size of the replication study. We can also invert
% equation \eqref{eq:res} to obtain the required replication sample size
% based on the specification of the minimum relative effect size $\dmin$
% to achieve replication success.
% This novel way of calculating the sample size requires the
% specification of the minimum relative effect size which can
% still be considered as acceptable.  Sample size calculations based on
% the two-trials rule can also be formulated in terms of the minimum
% relative effect size by inverting equation \eqref{eq:dSig}.
% We will report on a detailed comparison of the
% different approaches in future work. }

% \section*{Data and Software Availability}
% Data analyzed in this
% article and software are available in the R-package \texttt{ReplicationSuccess},
% which can be installed by running the following command in an R console:
% \texttt{install.packages("ReplicationSuccess", repos = \\"{http://R-Forge.R-project.org}"}).
% Further information on
% data preprocessing can be found on the corresponding help page
% (with the command \texttt{?RProjects}).

% \section*{Acknowledgments}
% Support by the Swiss
% National Science Foundation (Project \#~189295) is gratefully
% acknowledged. We acknowledge helpful and
% constructive comments by the Editor and a referee on an earlier version of this article.

\bibliographystyle{apalikedoiurl}
\bibliography{bibliography}