### 2022-11-30 abstract.tex
some fixes

% % Aus der Promotionsverordnung MNF §11 (4):
% % "Dem wissenschaftlichen Teil der Dissertation ist eine allgemein
% % verständliche kurze Zusammenfassung voranzustellen."

% \cleardoublepage
% \pdfbookmark{Zusammenfassung}{abstract_DE}

% { % need this extra grouping to avoid carry-over effects
% \begin{otherlanguage}{ngerman}
% \enlargethispage{4\baselineskip}
% \begin{center}\abstractnamefont
%   Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
%   incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis
%   nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
%   Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore
%   eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt
%   in culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum dolor
%   sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut
%   labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud
%   exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis
%   aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu
%   fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
%   culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum dolor sit
%   amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore
%   et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
%   ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor
%   in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
%   pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui
%   officia deserunt mollit anim id est laborum.

% \end{center}
% \end{otherlanguage}
% }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\cleardoublepage
\pdfbookmark{Abstract}{abstract_EN}

\enlargethispage{4\baselineskip}
\begin{center}\abstractnamefont
  \textbf{Abstract}
\end{center}

The credibility of a scientific finding is based on the fact that it can be
replicated when the same study is repeated with new subjects. However, various
failures to replicate major scientific findings in the social and life sciences
indicate that replicability is often lower than expected. This ``replication
crisis'' has led to several methodological reforms in the past decade, an
increased conduct of replication studies being one of them. Yet despite this
increase in replication studies, there is no consensus on a fundamental
question: How should replication success be defined statistically?
% Various methods already exist and various new ones have
% been proposed in response to the crisis. This thesis centres around a proposal
% based on the reverse-Bayes approach. The key idea is to reverse the traditional
% ``forward'' use of Bayes' theorem (``prior + likelihood $\rightarrow$
% posterior''), starting instead with a pre-specified posterior and deducing the
% corresponding prior (``posterior + likelihood $\rightarrow$
% prior''). % As such, the reverse-Bayes approach naturally fits to the replication
% % setting as it enables the mathematical formalization of scepticism regarding the
% % finding from the original study.
% This allows us to challenge the original finding by determining a sceptical
% prior for the underlying effect size such that the resulting posterior is no
% longer convincing. This prior then represents the position of a sceptic who
% remains unconvinced by the original study. Whether or not the sceptics' position
% is justified can then be assessed in light of the new data from the replication
% study. The larger the conflict between the replication data and the sceptic, the
% larger the degree of replication success. This procedure can be summarized in a
% single quantitative measure of replication success, termed the sceptical
% $p$-value.

The first and largest part of this thesis extends a recently proposed
reverse-Bayes method for quantifying replication success. The key idea is to
challenge the data from the original study with a sceptical prior so that the
resulting posterior no longer provides evidence for the finding from the study.
While the proposed method defined evidence through tail probabilities, the
extension developed here replaces them with Bayes
factors. % The sceptical prior is now
% determined such that the original finding is no longer convincing in terms of a
% Bayes factor. In contrast to tail probabilities,
We find that the extended method combines two notions of replicability: It
requires that original and replication study independently provide evidence for
an effect and ensures that their effect estimates are compatible. The extended
method also avoids a statistical paradox which the proposed method suffered
from.
% Bayes factors have a more
% natural interpretation and allow for direct quantification of evidence for one
% hypothesis versus another. As with the sceptical $p$-value, the procedure leads
% to a single measure quantifying the degree of replication success called the
% sceptical Bayes factor.
Another extension proposed in this thesis recalibrates the proposed method to
produce more appropriate inferences in terms of effect
size. % from the relative effect size
% perspective
The recalibration is chosen such that for original studies which were borderline
significant, replication success can only be achieved if the replication effect
estimate is larger than the original one. The recalibrated method has good
frequentist properties comparable to standard methods used in practice. The
third and last extension provides a framework for Bayesian design of replication
studies which allows for combining data from the original study with external
knowledge. External knowledge, such as data from other studies, is ubiquitous in
the replication setting and taking it into account can potentially lead to more
efficient designs compared to classical approaches.

The second part of the thesis is concerned with reverse-Bayes methodology. The
reverse-Bayes idea was first proposed in the 1950s, but it has mostly been
forgotten. To increase awareness and show potential use cases, reverse-Bayes
history and methods are summarized in a comprehensive review. The review
includes also some new results on connections between revere-Bayes methods and
meta-analysis. Finally, a commentary paper on a recently proposed reverse-Bayes
method draws connections to other reverse-Bayes methods.

The last part of the thesis revolves around research integrity issues in
methodological research. Questionable research practices, such as selective
reporting of results, are often seen as main cause for the low replicability in
applied research. These practices can similarly harm methodological research,
but are often not recognized. To raise awareness, an illustrative simulation
study is conducted in which it is shown how a novel method can easily be
presented as superior over established competitor methods if questionable
research practices are employed.

\textbf{Key words}: Bayesian inference, meta-science, replication studies

\newpage ~

\newpage

\begin{center}
  \vspace*{5cm}

  \textit{Dedicated to my mother Christa. \\
    You are deeply missed.}
\end{center}