### 2022-12-09 preface.tex
one-sided sceptical p

\section*{Preface}
\addcontentsline{toc}{chapter}{\texorpdfstring{\phantom{IV}Preface}{Preface}}

This thesis is submitted under the PhD program ``Epidemiology and
Biostatistics'' at the University of Zurich for the degree of Doctor of
Philosophy.
%(\url{http://www.epibiostat-phd.uzh.ch})
The research contained in this thesis was conducted between October 2019 and
December 2022. Financial support was provided by the Swiss National Science
Foundation through the project ``Reverse-Bayes design and analysis of
replication studies'' (project
\href{https://data.snf.ch/grants/grant/189295}{\#189295}) awarded to Leonhard
Held.

First and foremost I want to thank my supervisor Leo for giving me the
opportunity to do this PhD, for showing me an open-minded and pragmatic approach
to statistics, and for giving me the freedom to explore and develop into an
independent researcher. I also want to thank the other two members of my PhD
committee, Guido and Reinhard, for always providing excellent advice and being
great collaborators. Another thanks goes to Robert Matthews who also is a great
collaborator and without whom this PhD project would never exist as it was him
who resurrected the reverse-Bayes approach from the dead almost twenty years
ago. Furthermore, I want to thank Eric-Jan Wagenmakers for reviewing this thesis
and for giving me the opportunity to do a very interesting and productive six
months research stay in Amsterdam.

I also want to thank my friends and colleagues from the Epidemiology,
Biostatistics and Prevention Institute from the University of Zurich (in
alphabetical order): Ainesh, Alexandra, Annina, Charlotte, Babette, Bálint,
Beate, Beni, Dafne, Dominik, Eveline, Eva, Franscesca, Felix, Julia, Klaus,
Kelly, Lucas, Lisa, Luisa, Goscha, Manuela, Monika, Muriel, Manja, Maria,
Marielena, Martin, Mina, Nadja, Ruedi, Rachel, Sandra, Sona, Steffi, Tala,
Torsten, and Ulrike. Another thanks goes to my friends from the Department of
Psychological Methods from the University of Amsterdam (in alphabetical order):
Adam, Alexander, Alexandra, Alessandra, Alejandro, Angelika, Bruno, Don,
František, Frederik, Jason, Johnny, Joris, Jill, Julia, Lukas, Maarten,
Michelle, Nora, Omid, Quentin, René, Serjan, Suzanne, and Ting. I also want to
thank my good friends from outside of academia (in alphabetical order): Chronis,
Dani, Eleftheria, Flo, Fabi, Giuachin, Mirela, and Peter. A special thanks goes
to Ada. Finally, I thank my family: Beni, Christa, Elisabeth, Fabian, Harry,
Markus, Maja, and Noemi for their support.

\begin{flushright}
  \hfill Zürich, December 2022

  \hfill Samuel Pawel
\end{flushright}