### 2022-12-19 intro.Rnw
last polish of intro?

\chapter*{Introduction}
\phantomsection
\addcontentsline{toc}{chapter}{\texorpdfstring{\phantom{IV}Introduction}{Introduction}}

\hfill
\begin{tabular}{p{0.6\textwidth}}
  ``\emph{Confirmation comes from repetition}. Any attempt to avoid this statement leads at least to failure and more probably to destruction.'' \\
  \hfill J.W. \citet[p. 84]{Tukey1969}
\end{tabular}

\begin{wrapfigure}{r}{0.285\textwidth}
  \centering
  \includegraphics[width = 0.2\textwidth]{images/xkcdCrop.png}
  \caption*{Replication studies as illustrated by Randall Munroe
    (\url{https://xkcd.com/242}).}
\end{wrapfigure}

\section{Replication studies}

% How can we know if a finding from a scientific study is really true? For
% example, how can we know if the protective effect of a vaccine found in a study
% is real? The answer to such questions is of high importance to scientists,
% decision makers, and the general population. However, because any study result
% comes with uncertainty we can never know for sure.
Being able to trust scientific findings is essential for making evidence-based
decisions. But how can we know whether a scientific finding is trustworthy? For
instance, how can we know whether the protective effect of a vaccine found in a
study is indeed existent? One way to answer this question is through the use of
\emph{replication studies}.

A replication study (or simply replication) involves repeating the original
study as closely as possible but with new subjects. If the replication yields
similar results to the original study, this increases our confidence in the
original finding. However, if the replication study yields conflicting results,
this lowers our confidence in the original finding. %  As such, replication studies
% are an essential part of the scientific method. They provide a means for
% substantiating genuine research findings and refuting findings which occurred
% merely by chance.
The concept of replicability has long been a central part of the scientific
method. For instance, \citet[p. 13]{Fisher1935} noted ``\textit{no isolated
  experiment, however significant in itself, can suffice for the experimental
  demonstration of any natural phenomenon}''. For this reason, ``successful''
replication is typically a requirement for acceptance of newly proposed
scientific theories (\eg{} a new physical model). Yet, replication studies are
not only important for researchers, but also for decision makers and the general
population, as the implementation of policies based on scientific findings may
also depend on whether these findings can be replicated. For instance, for a
drug to be approved to the market, the United States food and drug
administration requires ``\textit{at least two adequate and well-controlled
  studies, each convincing on its own, to establish effectiveness}'' \citep[p.
3]{FDA1998}. The results from replication studies may hence have real-world
consequences. It is therefore important that the methodological aspects of
replication studies, such as their statistical design and analysis, are well
understood.

Despite their importance, the traditional academic system has made it
unattractive for researchers to conduct replication studies as publications,
citations, and grant money are typically easier to acquire by conducting novel
research. This is because until recently, most scientific journals hardly
published any replication studies since the emphasis was on ``innovation''
rather than on ``repetition'' \citep{Makel2012, Martin2017, Coiera2021}. As a
result, researchers often had no other choice than to focus their efforts on
producing novel and eye-catching research results in order to build successful
careers \citep{Binswanger2014, Moher2018}.

The perceived value of replication studies has changed over the past decade.
Earlier criticisms of low research standards \citep{Altman1994, Ioannidis2005}
were backed up by empirical evidence. For instance, pharmaceutical companies
reported surprisingly low replication rates from pre-clinical research
\citep{Begley2012} followed by later studies estimating that %\$28 billion
billions are wasted each year on flawed and non-replicable research in medicine
and the life sciences \citep{Chalmers2014, Freedman2015, Glasziou2018}.
Similarly, reports of fraud \citep{Wicherts2011} and questionable research
practices \citep{Wagenmakers2011, Simmons2011, John2012} sparked intense
discussion about the need for better research standards in psychology and the
social sciences. These discussions eventually culminated in large-scale
replication projects conducted by huge researcher consortia in fields such as
psychology \citep{Opensc2015, Klein2014, Klein2018, Protzko2020}, economics
\citep{Camerer2016}, the social sciences \citep{Camerer2018}, experimental
philosophy \citep{Cova2018}, and cancer biology \citep{Errington2021}.

Most of these large-scale replication projects confirmed what many researchers
had feared; carefully conducted replication studies often show less impressive
results than their original counterparts, and the replicability of research
findings is surprisingly low (how replicability can be defined precisely will
soon be discussed in more detail). This realization led many to declare science
as being in a ``replication crisis''. Debates arose about whether or not the
crisis really existed and who or what was to blame \citep{Gilbert2016,
  Amrhein2019}. Even the popular press became interested
\citep[\eg{}][]{Carey2015, Kovic2016, Achenbach2017, Devlin2018}, so that also
in the eyes of the public the credibility of science became seriously
threatened.

\begin{figure}[!htb]
<< "replication-time-graph", fig.height = 3.5 >>=
## ## Data gathered on October 18 2022
## searchTerm <- "Replication Study"
## searchDate <- as.Date(x = "2022-10-18", format = "%Y-%m-%d", tz = "GMT")
## searchDateFormat <- format(searchDate, format = "%B %d %Y")
## resultsAnytime <- 114000
## scholarDF <- data.frame(year = seq(2022, 1990, -1),
##                         results = c(4530, 11400, 13600, 12200, 11300, 10700,
##                                     9880, 9220, 8660, 8200, 7500, 6700, 6040,
##                                     5370, 4680, 3890, 3380, 3070, 2790, 2520,
##                                     2310, 2190, 2120, 2020, 1920, 1840, 1730,
##                                     1490, 1270, 1230, 1140, 1070, 970),
##                         study = c(347000, 459000, 1710000, 938000, 2360000,
##                                   675000, 3000000, 688000, 3150000, 512000,
##                                   3630000, 451000, 3650000, 396000, 3680000,
##                                   389000, 3890000, 375000, 3950000, 336000,
##                                   4160000, 364000, 4240000, 385000, 3470000,
##                                   377000, 3120000, 335000, 2620000, 376000,
##                                   2440000, 317000, 2220000))
## ## plot data
## ggplot(data = filter(scholarDF, year < 2022),
##        aes(x = year, y = results)) +
##     geom_segment(aes(xend = year, yend = 0), alpha = 0.2, size = 0.4) +
##     geom_point() +
##     ## scale_y_continuous(breaks = seq(0, 15000, 2500)) +
##     scale_x_continuous(breaks = seq(1990, 2020, 5)) +
##     ## scale_y_log10() +
##     labs(x = "Year", y = paste0('Number of results \nfor "', searchTerm, '"')) +
##     theme_bw() +
##     theme(panel.grid.minor = element_blank(),
##           panel.grid.major.x = element_blank()## ,
##           ## panel.grid.major.y = element_line(colour = "grey", linetype = "dashed")
##           )

WebofScienceDF <- data.frame(year = seq(2022, 1990, -1),
                             results = c(139, 144, 121, 137, 111, 80,
                                         89, 63, 68, 92, 63, 60, 39,
                                         39, 41, 30, 21, 20, 19, 25,
                                         15, 9, 13, 20, 19, 16, 14,
                                         8, 17, 13, 21, 5, 12),
                             study = c(235996, 272173, 245994, 230009, 207625,
                                       195012, 174800, 164635, 158745,
                                       184473, 138785, 128779, 116890, 111229,
                                       102231, 95143, 86691, 83132, 94432,
                                       79807, 70388, 53983, 54893, 53480,
                                       51744, 51898, 50070, 46826, 44591,
                                       43097, 42816, 43277, 43277)) %>%
    mutate(proportion = results/study,
           se = sqrt(proportion*(1 - proportion)/study))

## plot data
plotA <- ggplot(data = WebofScienceDF, aes(x = year, y = results)) +
    geom_segment(aes(xend = year, yend = 0), alpha = 0.2, linewidth = 0.5) +
    geom_point(size = 0.7) +
    scale_x_continuous(breaks = seq(1990, 2020, 5)) +
    lims(y = c(0, 160)) +
    labs(x = "Year", y = 'Number of results for "replication study"') +
    theme_bw() +
    theme(panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())
plotB <- ggplot(data = WebofScienceDF, aes(x = year, y = proportion)) +
    geom_segment(aes(y = proportion - 2*se, yend = proportion + 2*se, xend = year),
                 size = 0.5, alpha = 0.4) +
    geom_point(size = 0.7) +
    scale_x_continuous(breaks = seq(1990, 2020, 5), name = "Year") +
    scale_y_continuous(labels = scales::percent,
                       name = bquote("Proportion (" %+-% "2SE)")) +
    theme_bw() +
    theme(panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())

ggarrange(plotA, plotB, ncol = 2)
@
\caption{Yearly number of results for search term ``replication study'' on Web
  of Science (left), and number of results normalized by the number of results
  for the search term ``study'' (right). The search was conducted on 2 December
  2022.}
\label{fig:scholar}
\end{figure}

In the midst of this crisis, various reforms were implemented to prevent an
escalation of the situation \citep[for an overview see \eg{}][]{Munafo2017}. For
instance, some journals adopted the ``registered report'' format
\citep{Chambers2021} in which a study proposal is peer reviewed \emph{before}
the study is conducted, and which, if accepted, gives provisional publication
acceptance regardless of the study outcome. Similarly, digital infrastructure
platforms, such as Zenodo (\url{https://zenodo.org}) or the Open Science
Framework (\url{https://osf.io}), were created to facilitate pre-registration,
preprints, code and data sharing, all of which have substantially increased over
the last decade \citep{Kidwell2016, Nosek2018, Rawlinson2019}. The practice of
conducting replication studies has also gained popularity. Several academic
journals and funders are now explicitly promoting replication research
\citep{NWO2016, NSL2018, Nature2022}. Figure~\ref{fig:scholar} (left)
illustrates this trend by the yearly number of results for the search term
``replication study'' obtained from the search engine Web of Science. The
numbers have been rapidly growing, especially since the mid-2000s. This trend
could possibly be explained by the fact that more research results are published
each year but also when the numbers are normalized by the number of results for
the search term ``study'', the increasing trend is still visible.

\subsection{Statistical aspects of replication studies}
Despite the increased interest in replication studies, the research community
has not yet agreed on one important question: When is a replication study
considered successful? For this reason, replication researchers typically report
the results from different methods for assessing replication success. For
example, the \citet[p. 11]{Opensc2015} states ``[t]here is no single standard
for evaluating replication success. We evaluated [replicability] using
significance and P values, effect sizes, subjective assessments of replication
teams, and meta-analyses of effect sizes''. In the following, I will give an
overview about these and other methods which have been used in practice.

\begin{table}[!htb]
  \centering
  \caption{Study-level summary statistics for an original and a replication
    study. The cumulative distribution function of the standard normal
    distribution is denoted by $\Phi(\cdot)$, and the $1 - \alpha$ quantile of
    the standard normal distribution is denoted by
    $\Phi^{-1}(1 - \alpha) = \z{\alpha}$. Confidence intervals and
    \textit{p}-values for $H_{0}: \theta = 0$ are based on the assumption that
    an effect estimate $\that_{k}$ is normally distributed around the unknown
    effect size $\theta$ with (known) variance equal to the squared standard
    error $\sigma^{2}_{k}$ for $k \in \{o, r\}$.}
  \label{tab:meta}
  \begin{tabular}{lcc}
    \toprule
    & Original study & Replication study \\
    \midrule
    effect estimate & $\that_o$ & $\that_{r}$ \\
    standard error & $\sigma_{o}$ & $\sigma_{r}$ \\
    $(1 - \alpha)\times100\%$ confidence interval &
    % $[\that_{o} \pm \z{\alpha/2} \sigma_{o}]$ &
    % $[\that_{r} \pm \z{\alpha/2} \sigma_{r}]$ \\
    $[\that_{o} - \z{\alpha/2} \sigma_{o}, \, \that_{o} + \z{\alpha/2} \sigma_{o}]$ &
    $[\that_{r} - \z{\alpha/2} \sigma_{r}, \, \that_{r} + \z{\alpha/2} \sigma_{r}]$ \\
    $z$-value & $z_{o} = \that_{o}/\sigma_{o}$ & $z_{r} = \that_{r}/\sigma_{r}$ \\
    $p$-value (two-sided) & $p_{o} = 2\{1 - \Phi(|z_{o}|)\}$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    % One-sided $p$-value & $p_{o} = 1 - \Phi(|z_{o}|)$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    \bottomrule
  \end{tabular}
\end{table}

Most methods for analyzing replication studies can be formulated in terms of
study-level summary statistics as shown in Table~\ref{tab:meta}. All of these
are routinely reported in research articles, and if one of them is missing they
can typically be calculated from the others. Using summary statistics is also
often the only possible way for conducting the replication success analysis as
the raw data from the original study may not be available to the researchers
conducting the replication study. The most important statistics are the effect
estimates $\that_{o}$ and $\that_{r}$. They provide an estimate of the
underlying effect size $\theta$ which quantifies the true effect or association
of an intervention/exposure with an outcome variable. Typical effect sizes are
mean differences and correlations (for continuous outcomes), odds ratios (for
binary outcomes), or hazard ratios (for time-to-event outcomes). Depending on
the type of effect size, a transformation might be required so that the
assumption of approximately normally distributed effect estimates around the
unknown effect size $\theta$ (for large enough sample sizes) is justified. This
could be, for instance, the Fisher $z$-transformation for correlations or the
log-transformation for odds/hazard ratios \citep[chapter 11]{Cooper2019}. The
associated standard errors $\sigma_{o}$ and $\sigma_{r}$ represent the
statistical uncertainty of the estimates. Typically, the available standard
errors are only estimates of the true standard errors. However, in most parts of
this thesis it will be assumed that $\sigma_{o}$ and $\sigma_{r}$ correspond to
the true standard errors. This is the same assumption as in ordinary
fixed-effect/random-effects meta-analysis, and typically reasonable if the
sample size of the studies is not too small. Under the assumption of
(asymptotic) normality, confidence intervals for $\theta$ and $p$-values for
testing the null hypothesis $H_{0} \colon \theta = 0$ can be computed as shown
in Table~\ref{tab:meta}.

\begin{table}[!htb]
  \centering
  \caption{Statistical criteria for assessing replication success which have
    been used in practice \citep{Opensc2015, Camerer2016, Camerer2018, Cova2018,
      Errington2021}.}
  \label{tab:rsdef}
  \begin{tabular}{p{0.21\textwidth}p{0.7\textwidth}}
    \toprule
    Criterion type & Description \\
    \midrule
    Significance & \sloppy The original and replication $p$-values are smaller than
                   a threshold $\alpha$ and their effect estimates show
                   the same direction
                   (\mbox{$p_{o} < \alpha$}, \mbox{$p_{r} < \alpha$}, and
                   \mbox{$\sign(\that_{o}) = \sign(\that_{r})$}). Usually $\alpha = 5\%$.
    \\
                   & \\
    Meta-analytic significance & The meta-analytic $p$-value is smaller than a threshold
                                 $\alpha$ (\mbox{$p_{m} = 2\{1 - \Phi(|\that_{m}|/\sigma_{m})\} < \alpha$})
                                 where \mbox{$\that_{m} =
                                 (\that_{o}/\sigma_{o}^{2} + \that_{r}/\sigma^{2}_{r})\sigma^{2}_{m}$}
                                 is the  pooled effect estimate with standard error
                                 \mbox{$\sigma_{m} = (1/\sigma^{2}_{o} + 1/\sigma^{2}_{r})^{-1/2}$}.
                                 There are no established conventions regarding the level $\alpha$.
    \\
                   & \\
    Relative effect size & The effect estimate of the replication study goes in the same
                           direction as the original one and is at least $d_{\text{min}}$ times as large (\mbox{$d = \that_{r}/\that_{o} \geq d_{\text{min}}$}).
                           Usually $d_{\text{min}} = 1$. \\
                   & \\
    Confidence interval & \sloppy The replication effect estimate is contained in the
                          \mbox{$(1 - \alpha)\times100\%$} original confidence interval
                          ($\that_{r} \in [\that_{o} - z_{\alpha/2}\sigma_{o},
                          \that_{o} + z_{\alpha/2}\sigma_{o}]$).
                          Sometimes, also defined as the original effect estimate is
                          contained in the \mbox{$(1 - \alpha)\times100\%$} replication
                          confidence interval
                          (\mbox{$\that_{o} \in [\that_{r} - z_{\alpha/2}\sigma_{r},
                          \that_{r} + z_{\alpha/2}\sigma_{r}]$}).
                          Usually $\alpha = 5\%$. \\
                   & \\
    Prediction interval ($Q$-test) & The replication effect estimate is contained in its
                                     $(1 - \alpha)\times100\%$ prediction interval
                                     \mbox{($\that_{r} \in [\that_{o} - z_{\alpha/2}\surd(\sigma_{o}^{2}
                                     + \sigma_{r}^{2}), \that_{o} + z_{\alpha/2}\surd(\sigma_{o}^{2}
                                     + \sigma_{r}^{2})]$)}. Usually $\alpha = 5\%$.
                                     This criterion is equivalent to $p_{Q} \geq \alpha$ where
                                     $p_{Q}$ is the $p$-value from the meta-analytic $Q$-test
                                     $p_{Q} = 2\{1 - \Phi(\surd Q)\}$ with $Q$-statistic
                                     $Q = \sum_{i \in \{o, r\}} (\that_{i} - \that_{m})^{2}/\sigma_{i}^{2} = (\that_{o}- \that_{r})^{2}/(\sigma^{2}_{o} + \sigma^{2}_{r})$.\\
    \bottomrule
  \end{tabular}
\end{table}

Table~\ref{tab:rsdef} lists commonly used criteria for replication success in
terms of the summary statistics from Table~\ref{tab:meta}. The most popular
criterion is listed first and defines replication success by simultaneous
statistical significance of original and replication study along with their
effect estimates showing the same direction. This approach is also called the
``two-trials rule'' \citep{Senn2008} in drug regulation or ``vote-counting'' in
meta-analysis \citep{Cooper2019}. The criterion can similarly be defined through
one-sided $p$-values, so that the original and replication effect estimate are
not explicitly required to show the same direction as this is already taken into
account by the one-sided $p$-values. Some replication projects
\citep{Opensc2015, Errington2021} also defined replication success via
simultaneous non-significance of original and replication study
($p_{o} > \alpha$ and $p_{r} > \alpha$). However, with this definition
``replication success'' can almost always be achieved by conducting original and
replication study with very few samples so that the $p$-values are large. The
approach is also logically questionable as it could be seen as an instance of
the ``absence of evidence fallacy'' \citep{Altman1995} meaning the failure to
find evidence against the null hypothesis is erroneously interpreted as evidence
for the null hypothesis. Meta-analytic extensions of the significance approach
define replication success by significance of a combined effect
estimate.
% \footnote{The meta-analytic approach can also be given a Bayesian
% interpretation: A flat prior distribution for the underlying effect size
% $\theta$ is updated by the data from original and replication study.
% Replication success defined by a meta-analytic $p$-value being smaller than
% $\alpha$ is then equivalent to replication success via a Bayesian posterior
% probability $\Pr(\theta > 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$,
% respectively, $\Pr(\theta < 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$,
% depending on the direction of the combined estimate.}
Typically, the assumption of a common underlying effect size is seen as
reasonable so that fixed-effect meta-analysis is used for pooling.
Random-effects meta-analysis has primarily been used if more than one
replication study of the same original study are conducted since replicators are
often interested in understanding between-replication heterogeneity \citep[\eg{}
in][]{Klein2018}. In contrast to the ordinary significance criterion, it is less
established which level $\alpha$ should be used for thresholding the
meta-analytic $p$-value. Some replication projects have used the same level as
for the significance approach \citep{Opensc2015, Camerer2016, Errington2021},
whereas others have used a smaller level as the pooled estimate is based on both
data sets \citep[\eg{} $\alpha = 0.5\%$ in][]{Camerer2018}.

The remaining criteria in Table~\ref{tab:rsdef} put more emphasis on
compatibility in effect size between the original and the replication study. For
example, the relative effect estimate $d = \that_{r}/\that_{o}$ quantifies how
much the magnitude of the replication effect estimate changed compared to the
original, and the smaller $d$ is the smaller the degree of replication success.
Some projects also report a confidence interval for $d$ \citep{Camerer2016,
  Camerer2018}, whereas others ignore its uncertainty and make a binary cut at
$d_{\text{min}} = 1$ to define replication success \citep{Errington2021}. The
criteria based on confidence and prediction intervals define effect size
compatibility on an absolute scale. However, the criterion based on the original
confidence interval ignores the uncertainty from the replication, whereas the
criterion based on the replication confidence interval ignores the uncertainty
from the original study. The prediction interval criterion takes into account
both sources of uncertainty \citep{Patil2016}. Yet, declaring ``replication
success'' based on the prediction interval may be logically questionable due to
its connection to the meta-analytic \mbox{$Q$-test}. That is, if the $p$-value
from the $Q$-test is larger than $\alpha$ this is equivalent to the replication
effect estimate $\that_{r}$ being contained in its $(1- \alpha)\times 100\%$
prediction interval. The null hypothesis of this test is defined that the
underlying effect sizes of original and replication study are the same
($H_{0} \colon \theta_{o} = \theta_{r}$), so a rejection of this null hypothesis
corresponds to demonstrating replication failure and not replication success.
Interpreting a failure to reject the null hypothesis as evidence for it is again
an instance of the ``absence of evidence fallacy''
\citep{Hedges2019}. % To establish replication success a
% test based on the null hypothesis $H_{0} \colon \theta_{o} \neq \theta_{r}$
% would be required \citep{Hedges2019}.
As in the case of defining replication success by simultaneous non-significance
of original and replication study, the mismatch of the null hypothesis of the
prediction interval criterion results in the undesirable property that
``replication success'' can almost always be achieved if the sample size of the
studies is small enough, because the prediction interval becomes wider with
larger standard errors.

% The rapidly increasing conduct of replication studies has
% outpaced the development of methodology for their analysis which is perhaps the
% reason why many of these criteria have some logical flaws. For instance, the
% confidence interval criterion does only take into account the statistical
% variability of one of the two studies. As a result,

\subsubsection{Example ``Reproducibility Project: Cancer Biology''}
I will now illustrate the assessment of replication success on data from the
``Reproducibility Project: Cancer Biology'' \citep{Errington2021}. This
large-scale project attempted to replicate 53 landmark studies from the field of
cancer biology \citep[and would perhaps better be named ``Replication Project'',
see][for the distinction between ``replicabality'' and
``reproducibility'']{Goodman2016}. In the end, only 23 of the 53 studies could
be repeated due to various difficulties, such as, missing information from the
original studies or problems in conducting the experiments.
\citet{Errington2021} report that these experiments led to data on 158
quantitative effects. However, from the data which they provide only 132
quantitative effects are provided with original and replication standardized
mean difference effect estimates along with standard errors, and only they will
be used in the subsequent analyses.

\begin{figure}[!htb]
<< "RPCB-graph", fig.height = 3.75 >>=
za <- qnorm(p = 0.975)
rpcb <- read.csv("data/rpcb.csv") |>
    mutate(id = paste0("(", paper, ", ", experiment, ", ", effect, ")"),
           zo = smdo/so,
           zr = smdr/sr,
           c = so^2/sr^2,
           d = smdr/smdo,
           po2 = 2*(1 - pnorm(q = abs(zo))),
           pr2 = 2*(1 - pnorm(q = abs(zr))),
           sm = 1/sqrt(1/so^2 + 1/sr^2),
           smdm = (smdo/so^2 + smdr/sr^2)*sm^2,
           pm2 = 2*(1 - pnorm(q = abs(smdm/sm))),
           Q = (smdo - smdr)^2/(so^2 + sr^2),
           pQ = pchisq(q = Q, df = 1, lower.tail = FALSE),
           rsig = factor(pr2 <= 0.05, levels = c(FALSE, TRUE),
                         labels = c("italic(p[r]) > 0.05",
                                    "italic(p[r]) <= 0.05")))
## compute sceptical p-value (TODO fix problems with NA)
ind <- !is.na(rpcb$zo) & !is.na(rpcb$zr) & !is.na(rpcb$c)
rpcb$ps1 <- NA
rpcb$ps1[ind] <- pSceptical(zo = rpcb$zo[ind], zr = rpcb$zr[ind], c = rpcb$c[ind],
                            alternative = "one.sided", type = "nominal")

zoom <- c(-2, 12)
alphazoom <- 0.6
ltyzoom <- 3
rpcbWithoutNA <- na.omit(rpcb)
cols <- c("#1A1A1A80", "#1B9E77")
names(cols) <- levels(rpcbWithoutNA$rsig)
plotA <- ggplot(data = rpcbWithoutNA, aes(x = smdo, y = smdr)) +
    annotate(geom = "segment", x = zoom[1], xend = zoom[2],
             y = zoom[1], yend = zoom[1], alpha = alphazoom,
             lty = ltyzoom) +
    annotate(geom = "segment", x = zoom[1], xend = zoom[2],
             y = zoom[2], yend = zoom[2], alpha = alphazoom,
             lty = ltyzoom) +
    annotate(geom = "segment", x = zoom[1], xend = zoom[1],
             y = zoom[1], yend = zoom[2], alpha = alphazoom,
             lty = ltyzoom) +
    annotate(geom = "segment", x = zoom[2], xend = zoom[2],
             y = zoom[1], yend = zoom[2], alpha = alphazoom,
             lty = ltyzoom) +
    geom_abline(intercept = 0, slope = 1, alpha = 0.3) +
    geom_hline(yintercept = 0, lty = 2, alpha = 0.9, size = 0.4) +
    geom_vline(xintercept = 0, lty = 2, alpha = 0.9, size = 0.4) +
    geom_point(aes(fill = rsig), shape = 21, alpha = 0.9) +
    geom_rug(aes(color = rsig), size = 0.3, show.legend = FALSE) +
    labs(x = "Original effect estimate (SMD)",
         y = "Replication effect estimate (SMD)",
         fill = "") +
    coord_fixed(xlim = c(-5, 30), ylim = c(-5, 30)) +
    scale_fill_manual(values = cols, labels = scales::parse_format()) +
    scale_color_manual(values = cols) +
    theme_bw() +
    theme(panel.grid = element_blank(),
          legend.position = c(0.5, 0.835))

plotB <- ggplot(data = rpcbWithoutNA, aes(x = smdo, y = smdr)) +
    geom_abline(intercept = 0, slope = 1, alpha = 0.3) +
    geom_hline(yintercept = 0, lty = 2, alpha = 0.9, size = 0.4) +
    geom_vline(xintercept = 0, lty = 2, alpha = 0.9, size = 0.4) +
    geom_point(aes(fill = rsig), shape = 21, alpha = 0.9,
               show.legend = FALSE) +
    geom_rug(aes(color = rsig), size = 0.3, show.legend = FALSE) +
    labs(x = "Original effect estimate (SMD)",
         y = "Replication effect estimate (SMD)",
         fill = "") +
    coord_fixed(xlim = zoom, ylim = zoom) +
    scale_fill_manual(values = cols, labels = scales::parse_format()) +
    scale_color_manual(values = cols) +
    theme_bw() +
    theme(panel.grid = element_blank())
ggpubr::ggarrange(plotA, plotB, ncol = 2, align = "v")
@
\caption{Results for 132 effects from the ``Reproducibility Project: Cancer
  Biology'' \citep{Errington2021} for which effect estimates and standard errors
  are available on standardized mean difference (SMD) scale. The right plot
  shows a zoomed-in view of the dotted area in the left plot. The
  \textit{p}-values are recomputed using a normal approximation. Two study pairs
  with original effect estimate larger than 80 are not shown.}
\label{fig:rpcb}
\end{figure}

Figure~\ref{fig:rpcb} shows the original versus the replication effect estimate
(on standardized mean difference scale) with the color indicating whether the
replication study was statistically significant at the 5\% level (two-sided). As
in most other replication projects, the majority of the replications show
smaller effect estimates compared to their original counterparts (mean relative
effect estimate $\bar{d} = \Sexpr{round(mean(rpcbWithoutNA[,"d"]), 2)}$). Many
of the replications also fail to achieve statistical significance at the 5\%
level. Specifically, from the
\Sexpr{sum(rpcbWithoutNA$po2 < 0.05)} effects which were significant in the original study only \Sexpr{ sum(rpcbWithoutNA$po2
  < 0.05 & rpcbWithoutNA$pr2 < 0.05 & rpcbWithoutNA$d > 0)} were also
significant in the replication study (in the same direction).

\begin{figure}[!htb]
<< "RPCB-table1", results = "asis", fig.height = 6.5 >>=
## show number of successes
alpha <- 0.05
dmin <- 1
rpcbUpset<- rpcb |>
    filter(!is.na(po2), !is.na(pr2), !is.na(smdo), !is.na(smdr),
           !is.na(so), !is.na(sr)) |>
    mutate("Significance" = as.numeric(po2 < alpha & pr2 < alpha & sign(smdo) == sign(smdr)),
           "Meta analytic \nsignificance" = as.numeric(pm2 < alpha),
           "Relative \neffect size" = as.numeric(d >= dmin),
           "Prediction \ninterval" = as.numeric(pQ > alpha))
nsig <- sum(rpcbUpset[,"Significance"], na.rm = TRUE)
nma <- sum(rpcbUpset[,"Meta analytic \nsignificance"], na.rm = TRUE)
nres <- sum(rpcbUpset[,"Relative \neffect size"], na.rm = TRUE)
npi <- sum(rpcbUpset[,"Prediction \ninterval"], na.rm = TRUE)

upset(rpcbUpset, sets = c("Significance", "Meta analytic \nsignificance",
                          "Relative \neffect size", "Prediction \ninterval"),
      order.by = "freq", mainbar.y.label = "Successful replications (combined criteria)",
      sets.x.label = "Successful replications \n(single criterion)", #, empty.intersections = "on")
      queries = list(list(query = intersects,
                          params = c("Significance",
                                     "Meta analytic \nsignificance",
                                     "Relative \neffect size",
                                     "Prediction \ninterval"),
                          color = "#1b9e77", active = TRUE)))
@
\caption{UpSet plot for data on 132 effects from the ``Reproducibility Project:
  Cancer Biology'' \citep{Errington2021} for which effect estimates and standard
  errors are available on the standardized mean difference scale. Shown are the
  number of replication successes according to the different criteria from
  Table~\ref{tab:rsdef} and their combinations. A threshold of
  $\alpha = \Sexpr{round(100*alpha, 2)}\%$ and a relative effect size threshold
  of $d_{\text{min}} = \Sexpr{dmin}$ are used for replication success.}
\label{fig:upset}
\end{figure}

Figure~\ref{fig:upset} shows how many of the replications are successful
according to the criteria from Table~\ref{tab:rsdef} (except the confidence
interval criteria since they do not take into account the uncertainty from both
studies). For the total 132 replications, most successes occur for the
meta-analytic significance (\Sexpr{nma}) and the prediction interval criteria
(\Sexpr{npi}), followed by significance (\Sexpr{nsig}), and relative effect size
(\Sexpr{nres}). However, looking at the combinations of the criteria, only in
two replications are all criteria satisfied simultaneously (green).

A detailed view for a subset of the data from the project is given in
Table~\ref{tab:RPCB}. The subset is randomly chosen such that from every
original study one effect and one replication of that effect is included. For
the three replications at the bottom of the table (\#18, \#19, and \#20), all
criteria indicate replication failure. For the other replications, the
conclusions are more ambiguous as in no single case are all criteria met
simultaneously. For instance, the first three replications satisfy the
significance criterion ($p_{o} < 0.05$ and $p_{r} < 0.05$), the meta analytic
significance criterion ($p_{m} < 0.05$) and the $Q$-test/prediction interval
criterion ($p_{Q} > 0.05$), yet the replication effect estimate is smaller than
the original one so does not satisfy the relative effect size criterion
($d < 1$). The fourth replication, on the other hand, satisfies the relative
effect size criterion ($d \geq 1$) but fails to satisfy the significance
criterion as the original study was not significant ($p_{o} > 0.05$).
% For example, replication \#17 is a clear
% failure with respect to the relative effect estimate ($d < 1$) and the
% significance criteria ($p_{o} > 0.05$ and $p_{r} > 0.05$), yet the $Q$-test does
% not indicate evidence for inconsistency of the two effect estimates
% ($p_{Q} > 0.05$).

Taken together, this analysis demonstrates that conclusions based on commonly
used replication success criteria often differ. It is not always clear-cut
whether or not a replication is successful. Reducing replicability to a single
criterion without mentioning these difficulties, as often done by the popular
press \citep[\eg{} ``more than half of the findings did not hold up when
retested'' in][]{Carey2015}, is a simplification of the matter.

\begin{landscape}
\begin{table}
  \centering
  \caption{Subset of results from the ``Reproducibility Project: Cancer
    Biology'' \citep{Errington2021}. (P, X, E) denotes effect number E from
    experiment number X, from original paper number P. Shown are original
    $\that_{o}$, replication $\that_{r}$, and pooled effect estimate $\that_{m}$
    with 95\% confidence intervals, variance ratio
    $c = \sigma^{2}_{o}/\sigma^{2}_{r}$, relative effect estimate
    $d = \that_{r}/\that_{o}$, original \textit{p}-value $p_{o}$, replication
    \textit{p}-value $p_{r}$, meta-analytic \textit{p}-value $p_{m}$, $Q$-test
    \textit{p}-value $p_{Q}$, and one-sided sceptical \textit{p}-value
    $p_{\mathrm{S}}$. A replication success (green-bold) threshold of $\alpha = 5\%$
    is used for two-sided \textit{p}-values, $\alpha = 2.5\%$ for one-sided
    $p$-values, and $d_{\text{min}} = 1$ for the relative effect estimate.}
  \label{tab:RPCB}
\resizebox{1.02\linewidth}{!}{%
<< "RPCB-table2", results = "asis" >>=
## function to format CI
ciFormat. <- function(lower, upper, dollar = FALSE, digits = 2) {
    if (dollar) x <- "$" else x <- ""
    paste0("(", x, format(round(lower, digits), nsmall = digits), x,
           " to ", x, format(round(upper, digits), nsmall = digits), x, ")")
}
ciFormat <- Vectorize(FUN = ciFormat.)

## function to format p-value
pFormat. <- function(p, digits = 2, dollar = FALSE, phantom = FALSE, font = "mathrm") {
    if (is.na(p)) {
        out <- NA
    } else {
        if (p > 0.0001) {
            if (p < 10^(-digits+1)) {
                out <- as.character(signif(p, digits = digits))
            } else {
                out <- format(round(p, digits), nsmall = digits)
            }
            if (phantom) out <- paste0("\\phantom{<~}", out)
        } else {
            out <- "< 0.0001"
        }
        if (dollar) {
            out <- paste0("$\\", font, "{", out, "}$")
        }
    }
    return(out)
}
pFormat <- Vectorize(FUN = pFormat.)

## function to format (relative) effect estimate
dFormat. <- function(x, digits = 2, phantom = FALSE, dollar = FALSE, font = "mathrm") {
    if (is.na(x)) {
        out <- NA
    } else {
        out <- format(round(x, digits), nsmall = digits)
        if (phantom && (x >= 0)) {
           out <- paste0("\\phantom{-}", out)
        }
        if (dollar) {
            out <- paste0("$\\", font, "{", out, "}$")
            }
    }
    return(out)
}
dFormat <- Vectorize(FUN = dFormat.)

## function to color code numerical value in LaTeX table based on threshold
tableFormat. <- function(x, outFun = pFormat, threshold = 0.05, col1 = "dark2green",
                         col2 = "dark2orange", font1 = "mathbf", font2 = "mathrm") {
    if (is.na(x)) {
        out <- NA
    } else {
        if (x < threshold) {
            col <- col1
            font <- font1
        } else {
            col <- col2
            font <- font2
        }
        out <- paste0("\\textcolor{", col, "}{",
                      outFun(x, dollar = TRUE, phantom = TRUE, font = font),
                      "}")
    }
    return(out)
}
tableFormat <- Vectorize(FUN = tableFormat., vectorize.args = "x")

## randomly select a subset with one effect per paper
set.seed(7)
rpcbTable <- rpcb |>
    na.omit() |>
    group_by(paper) |>
    sample_n(size = 1) |>
    ungroup() |>
    arrange(ps1) |>
    mutate(smdoFormat = signif(smdo, 3),
           smdrFormat = signif(smdr, 3),
           smdmFormat = signif(smdm, 3),
           dFormat = tableFormat(x = d, outFun = dFormat,
                                 threshold = 1, col1 = "dark2orange", font1 = "mathrm",
                                  col2 = "dark2green", font2 = "mathbf"),
           cioL = smdo - za*so,
           cioU = smdo + za*so,
           cirL = smdr - za*sr,
           cirU = smdr + za*sr,
           cimL = smdm - za*sm,
           cimU = smdm + za*sm,,
           po2Format = tableFormat(x = po2),
           pr2Format = tableFormat(x = pr2),
           pm2Format = tableFormat(x = pm2),
           ps1Format = tableFormat(x = ps1, threshold = 0.025),
           pQFormat = tableFormat(x = pQ, col1 = "dark2orange", font1 = "mathrm",
                                  col2 = "dark2green", font2 = "mathbf"),
           cioFormat = ciFormat(lower = cioL, upper = cioU, dollar = TRUE),
           cirFormat = ciFormat(lower = cirL, upper = cirU, dollar = TRUE),
           cimFormat = ciFormat(lower = cimL, upper = cimU), dollar = TRUE) |>
    select(id,
           smdoFormat, cioFormat,
           smdrFormat, cirFormat,
           smdmFormat, cimFormat,
           c, dFormat,
           po2Format, pr2Format, pm2Format, pQFormat, ps1Format
           )
rpcbXtable <- xtable(rpcbTable)
colnames(rpcbXtable) <- c("(P, X, E)",
                          "$\\that_o$", "(95\\% CI)",
                          "$\\that_r$", "(95\\% CI)",
                          "$\\that_m$", "(95\\% CI)",
                          "$c$",
                          "$d$",
                          "$p_o$",
                          "$p_r$",
                          "$p_m$",
                          "$p_Q$",
                          "$p_{\\text{S}}$"
                          )

colnames(rpcbXtable) <- paste0("\\multicolumn{1}{c}{", colnames(rpcbXtable), "}")
print(rpcbXtable, booktabs = TRUE, floating = FALSE, include.rownames = TRUE,
      sanitize.text.function = function(x){x},
      math.style.negative = getOption("xtable.math.style.negative", TRUE),
      size = "small")
@
%
}
\end{table}

\end{landscape}

\subsection{Reverse-Bayes assessment of replication studies}
In response to the lack of a standard criterion for replication success, various
methods have been proposed \citep[among others]{Verhagen2014, Simonsohn2015,
  Anderson2016, Patil2016, Johnson2016, Etz2016, vanAert2017, Ly2018, Harms2019,
  Hedges2019, Mathur2020, Held2020, Pawel2020, Bonett2020}. The focus of this
thesis is to refine and extend the proposal from \citet{Held2020} which combines
\emph{reverse-Bayes inference} and \emph{Bayesian model criticism} in a method
for assessing replication success. In the following, I will summarize the method
and its technical underpinnings.

\subsubsection{Reverse-Bayes inference}

Bayesian inference is an approach to statistical inference where Bayes' theorem
is used to make probability statements about unknown parameters based on the
observed data. The central quantity for doing so is the distribution of the
parameters conditional on the data, called the \emph{posterior distribution}. It
can be obtained from Bayes' theorem,
\begin{align*}
  f(\theta \given \text{data})
  = f(\theta) \times \frac{f(\text{data} \given \theta)}{f(\text{data})},
\end{align*}
meaning that the \emph{prior distribution} for the parameter $\theta$ with
probability density/mass function $f(\theta)$ is multiplied by the (normalized)
likelihood of the data, also know as \emph{Bayesian updating}. Parameter values
which increase the likelihood of the data become more likely \emph{a posteriori}
but they are also weighted by their \emph{a priori} plausibility through the
prior. As such, Bayesian inference provides a formal way for combining
information from the data at hand with external knowledge encoded in the prior.

Many consider the prior to be also the weak point of Bayesian inference since it
is unclear how it should be specified in the absence of external knowledge. The
\emph{reverse-Bayes} approach, first proposed by \citet{Good1950}, is one way of
dealing with this issue. The idea is to reverse Bayes' theorem as
\begin{align*}
  f(\theta) = f(\theta \given \text{data}) \times
  \frac{f(\text{data})}{f(\text{data} \given \theta)}
\end{align*}
and instead ``downdate'' a posterior with the observed data. So, in contrast to
the conventional ``forward-Bayes'' approach where we start with a prior, update
it with the data, and end up with a posterior, the reverse-Bayes approach starts
with the posterior and ends up with the prior. For example, the posterior may be
specified to indicate evidence for/against a particular hypothesis under
investigation. Reverse-Bayes inference then revolves around the question of
whether the resulting prior is plausible in light of external knowledge, and if
so, this indicates support for the specified posterior.

To illustrate reverse-Bayes inference, let us return to the replication setting.
Assume we want to conduct inference about the unknown effect size $\theta$ based
on the effect estimate from the original study $\that_{o}$ and its standard
error $\sigma_{o}$. We will assume that $\that_{o}$ is normally distributed
around the unknown effect size $\theta$ with (known) variance equal to its
squared standard error $\sigma^{2}_{o}$, here and henceforth denoted by
$\that_{o} \given \theta \sim \Nor(\theta, \sigma^{2}_{o})$. Furthermore, we
specify a zero-mean normal prior with variance $\tau^{2}$ for the effect size
$\theta$, representing the position of a sceptic who does not believe in the
presence of a genuine effect. The ``stubbornness'' of the sceptic is determined
by how small the variance $\tau^{2}$ is chosen (the smaller $\tau^{2}$, the more
scepticism). Combining the sceptical prior with the likelihood produces a
posterior which is again normal
$\theta \given \that_{o}, \sigma_{o} \sim \Nor(\mu_{\text{post}}, \sigma^{2}_{\text{post}})$
with mean and variance
\begin{align*}
  &\mu_{\text{post}} = \frac{\that_{o}}{1 + \sigma^{2}_{o}/\tau^{2}}&
  &\text{and}&
  &\sigma^{2}_{\text{post}} = \frac{1}{1/\sigma^{2}_{o} + 1/\tau^{2}}.&
\end{align*}
The associated $(1- \alpha)\times 100\%$ highest posterior density credible
interval has limits
\begin{align}
  \mu_{\text{post}} \pm z_{\alpha/2} \, \sigma_{\text{post}}
  \label{eq:postCI}
\end{align}
and if this credible interval excludes parameter values smaller/larger than zero
(depending on the orientation of the effect size) this may be interpreted as
evidence for a genuine effect found in the original study. There exist also
other definitions of statistical evidence and in subsequent chapters of this
thesis (Paper~\ref{PaperII} and Paper~\ref{PaperIV}) we will extend this
reverse-Bayes procedure to employ alternative measures of evidence, such as
Bayes factors.

Depending on how large the prior variance $\tau^{2}$ is chosen, the posterior
credible interval~\eqref{eq:postCI} will either include or exclude zero.
Different researchers may have different degrees of scepticism and may thus
choose different prior variances $\tau^{2}$. As a default choice,
\citet{Held2020} proposed to use the reverse-Bayes approach from
\citet{Matthews2001b}, that is, to determine the \emph{sufficiently sceptical
  prior variance} $\tau^{2}_{\alpha}$ so that the appropriate limit of the
$(1 - \alpha)\times 100\%$ credible interval is just fixed to zero. The resulting prior then
represents the beliefs of a sceptic who is just stubborn enough to not find the
original study convincing at level $\alpha$.

\begin{figure}[!htb]
<< "RBexample", fig.height = 3 >>=
## show reverse-Bayes for replications for a study and its three replications
exDat <- rpcb |>
    filter(id == "(9, 2, 5)")
to <- unique(exDat$smdo)
so <- unique(exDat$so)
zo <- to/so
tr <- exDat$smdr
sr <- exDat$sr

## do reverse-Bayes calculations
alpha <- 0.05
za <- qnorm(p = 1 - alpha/2)
sprior <- sqrt(so^2/(zo^2/za^2 - 1))
sposterior <- sqrt(1/(1/sprior^2 + 1/so^2))
mposterior <- sposterior^2*to/so^2
plotDF <- data.frame(t = c(to, tr, 0, mposterior),
                     s = c(so, sr, sprior, sposterior),
                     type = c("Original", rep("Replication", 3), "Prior",
                              "Posterior"))
plotDF$type <- factor(plotDF$type, levels = c("Original", "Posterior", "Prior",
                                              "Replication"))

## compute prior density
xprior <- seq(-3*sprior, 3*sprior, length.out = 200)
dprior <- dnorm(x = xprior, mean = 0, sd = sprior)
dpriorScaled <- 0.75*dprior#*max(dprior)

## compute posterior density
xpost <- seq(mposterior - 3*sposterior, mposterior + 3*sposterior,
             length.out = 200)
dpost <- dnorm(x = xpost, mean = mposterior, sd = sposterior)
dpostScaled <- 0.75*dpost#*max(dpost)

## prior data conflict assessment
pbox <- pchisq(q = tr^2/(sr^2 + sprior^2), df = 1, lower.tail = FALSE)
pboxFormat <- pFormat(p = pbox)
pboxLabel <- paste("italic(p)['Box'] == ", pboxFormat)
plotDF$Label <- c(NA, pboxLabel, NA, NA)
dodgewidth <- 0.5

ggplot(data = plotDF, aes(x = type, y = t)) +
    geom_hline(yintercept = 0, lty = 2, alpha = 0.3) +
    geom_pointrange(aes(ymin = t - za*s, ymax = t + za*s, col = type),
                    fatten = 3,
                    position = position_dodge2(width = dodgewidth),
                    show.legend = FALSE) +
    annotate("segment", x = 1.81, y = -0.5, xend = 1.97, yend = -0.05,
             alpha = 0.6,
             arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
    annotate("text", x = 1.85, y = -0.6, size = 3, alpha = 0.6,
             label = paste0(round(100*alpha/2, 2), "% quantile fixed at zero")) +
    annotate("segment", x = 2.8, y = -0.5, xend = 2.96, yend = -0.05,
             alpha = 0.6,
             arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
    annotate("text", x = 2.75, y = -0.6, size = 3, alpha = 0.6,
             label = "fixed at zero") +
    annotate("segment", x = 1.6, y = 2.25, xend = 2.7, yend = 2.25,
             alpha = 0.6,
             arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
    annotate("text", x = 2.1, y = 2.45, size = 3, alpha = 0.6,
             label = "reverse-Bayes analysis") +
    annotate("segment", x = 3, y = 2.25, xend = 4, yend = 2.25,
             alpha = 0.6,
             arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
    annotate("text", x = 3.45, y = 2.45, size = 3, alpha = 0.6,
             label = "assess prior-data conflict") +
    annotate("path", x = 3 + dpriorScaled, y = xprior, color = 2, alpha = 0.5) +
    annotate("path", x = 2 + dpostScaled, y = xpost, color = 4, alpha = 0.5) +
    geom_text(aes(label = Label, y = t - za*s*0.9),
              position = position_dodge2(width = dodgewidth),
              parse = TRUE, size = 2.75, alpha = 0.8, hjust = -0.08) +
    labs(x = NULL, y = "Effect size") +
    scale_color_manual(values = c("Original" = 1, "Replication" = 1,
                                  "Posterior" = 4, "Prior" = 2)) +
    theme_bw() +
    theme(panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank(),
          axis.text.x = element_text(size = 11))
@
\caption{Illustration of reverse-Bayes assessment of replication success using
  data from the original study (Paper 9, Experiment 2, Effect 5) and its three
  replication studies from the ``Reproducibility Project: Cancer Biology''
  \citep{Errington2021}. Shown are effect estimates and prior/posterior means
  with \Sexpr{100*(1-alpha)}\% confidence/credible interval. The original finding is
  challenged with a sceptical prior, sufficiently concentrated around zero so
  that the resulting posterior is no longer convincing at level
  $\alpha = \Sexpr{100*(alpha)}\%$. Prior predictive \textit{p}-values $p_{\text{Box}}$ are
  computed for quantifying prior-data conflict.}
\label{fig:RBillustration}
\end{figure}

Figure~\ref{fig:RBillustration} illustrates the derivation of the sufficiently
sceptical prior variance for an original study included in the ``Reproducibility
Project: Cancer Biology'' with standardized mean difference effect estimate
$\that_{o} = \Sexpr{round(to, 2)}$ and standard error
$\sigma_{o} = \Sexpr{round(so, 2)}$. We see that the sufficiently sceptical
prior with variance $\tau_{\alpha}^{2} = \Sexpr{round(sprior, 2)}^{2}$ produces
a posterior with $1 - \alpha = \Sexpr{round(100*(1 - alpha), 2)}\%$ credible
interval fixed to zero, so that the original finding is rendered no longer
convincing at level $\alpha = \Sexpr{round(100*alpha, 2)}$\%.

\citet{Held2019a} showed that the sufficiently sceptical prior variance
$\tau^{2}_{\alpha}$ for a level $\alpha$ is available in closed-form
\begin{align}
  \tau^{2}_{\alpha} =
  \begin{cases}
    \dfrac{\sigma^{2}_{o}}{\left(z_{o}^{2}/z_{\alpha/2}^{2}\right) - 1}
    & \text{if} ~ z_{o}^{2} > z_{\alpha/2}^{2}, \\
    \text{undefined} & \text{else.}
  \end{cases}
  \label{eq:tau2alpha}
\end{align}
From~\eqref{eq:tau2alpha} we see that convincing original studies (those with
large absolute $z$-values $|z_{o}|$) require smaller sufficiently sceptical
prior variances to render the posterior no longer convincing for the same level
$\alpha$. Conversely, if the original study is not convincing enough (if
$|z_{o}| \leq z_{\alpha/2}$) the sufficiently sceptical prior variance is
undefined meaning that the data provide so little evidence that no scepticism is
required to make them unconvincing.

In this way the reverse-Bayes approach based on sceptical priors can be used to
formally challenge the finding from an original study. However, once this
sceptical prior is determined, the question becomes whether it is plausible in
light of external data. A natural candidate for answering the question is data
from a replication study. In the following, I will show how Bayesian model
criticism can be used for doing so.

\subsubsection{Bayesian model criticism}
Model criticism describes the assessment of compatibility between observed data
and their assumed statistical model. If incompatibility is diagnosed, this
alarms the data analyst that inferences based on the model may be invalid and
modifications may be required. A formal framework for Bayesian model criticism
was first introduced by \citet{Box1980}. To understand whether a Bayesian model
M consisting of a joint distribution for parameter $\theta$ and data is
adequate, Box gave the following fundamental decomposition of the joint
distribution
\begin{align*}
  f(\theta, \text{data} \given \text{M})
  % &= f(\theta \given \text{M}) \times f(\text{data} \given \theta, \text{M})
  % \label{eq:decomp1} \\
  &= f(\theta \given \text{data}, \text{M}) \times f(\text{data} \given \text{M}).
    % \label{eq:decomp2}.
\end{align*}
% He noted from the first decomposition~\eqref{eq:decomp1} that the model M may
% only be regarded as adequate if current belief about
He argued that inferences based on the left factor, the posterior distribution
$f(\theta \given \text{data}, \text{M})$, should only be trusted if the right
factor, the prior predictive distribution
\begin{align*}
  f(\text{data} \given \text{M})
  = \int f(\text{data} \given \theta, \text{M}) \,
  f(\theta \given \text{M}) \, \text{d}\theta
\end{align*}
is compatible with the observed data. If the model M was indeed adequate, the
empirical distribution of the observed data should be close to its predictive
distribution under the model M. On the other hand, if the empirical distribution
differed from the predictive distribution, this would imply that model M is
inadequate due to misspecification of the likelihood
$f(\text{data} \given \theta, \text{M})$ and/or misspecification of the prior
$f(\theta \given \text{M})$.

Based on these observations, Box proposed two general approaches for conducting
Bayesian model criticism. First, the predictive density of the observed data (or
the predictive density of a ``checking function'' applied to the observed data)
can be compared to its reference distribution via a \emph{prior predictive
  $p$-value}
\begin{align}
  p_{\text{Box}}
  % &= \Pr\left\{f(\text{data} \given \text{M}) <
  % f(\text{data} = \text{observed data} \given \text{M}) \given \text{M}\right\} \\
  & = \int_{\mathcal{X}} \,
    f(\text{data} = x \given \text{M}) \, \text{d}x
  \label{eq:pppval}
\end{align}
with $\mathcal{X} = \{x : f(\text{data} = x \given \text{M}) <
f(\text{data} = \text{observed data} \given \text{M})\}$.
The $p$-value $p_{\text{Box}}$ is the probability of obtaining data with lower
predictive density (``more surprising'' data) than the observed data, and the
lower $p_{\text{Box}}$, the more incompatibility between the observed data and
the assumed model M. This approach was used by \citet{Held2020}.
% The literature on Bayesian model criticism following \citet{Box1980} also
% mostly focused on prior predictive $p$-values. However, what has been mostly
% ignored is that Box also mentioned an alternative approach.
However, Box also mentioned a second approach which has mostly been forgotten.
If a second ``benchmarking'' model $\text{M}_{2}$ alternative to the model under
investigation $\text{M}_{1}$ is available, Box proposed that the \emph{prior
  predictive ratio}
\begin{align*}
  \text{PPR}_{\text{Box}}
  = \frac{f(\text{data} = \text{observed data} \given \text{M}_{1})}{
  f(\text{data} = \text{observed data} \given \text{M}_{2})},
\end{align*}
the ratio of predictive densities from the observed data under both models,
could be used to judge the relative adequacy of model $\text{M}_{1}$. Again, the
lower the prior predictive ratio $\text{PPR}_{\text{Box}}$, the less compatible
the observed data with the model $\text{M}_{1}$. Bayesian model criticism
approaches based on prior predictive ratios will be used in subsequent chapters
of this thesis (Paper~\ref{PaperII} and Paper~\ref{PaperIV}).
% \footnote{Some readers may have noted that the prior predictive ratio is also
% the Bayes factor \citep{Jeffreys1961, Good1958} contrasting model
% $\text{M}_{1}$ to $\text{M}_{2}$. The prior predictive ratio model criticism
% approach is therefore particularly useful in combination with reverse-Bayes
% procedures based on Bayes factors, as will be demonstrated later (Paper~\ref{PaperII} and
% IV).}

We now return to the replication setting. Having obtained a sceptical prior
$\theta \sim \Nor(0, \tau^{2}_{\alpha})$ with sufficiently sceptical prior
variance $\tau^{2}_{\alpha}$ from~\eqref{eq:tau2alpha}, the aim is to assess its
adequacy in light of the data from a replication study. If we are able to show
that the prior is inadequate, this would demonstrate that scepticism regarding
the original finding is unjustified and that the original study provided
evidence for a genuine effect. Under the assumption of a normal likelihood for
the replication effect estimate, \ie{}
$\that_{r} \given \theta \sim \Nor(\theta, \sigma^{2}_{r})$, the prior
predictive distribution is given by
$\that_{r} \given \theta \sim \Nor(0, \sigma^{2}_{r} + \tau^{2}_{\alpha})$. As
the prior predictive distribution is symmetric around zero, the prior predictive
$p$-value~\eqref{eq:pppval} is
\begin{align}
  p_{\text{Box}}
  = 2\left\{1 - \Phi\left(\frac{|\that_{r}|}{\sqrt{\sigma^{2}_{r} +
  \tau^{2}_{\alpha}}}\right)\right\}.
  \label{eq:pboxNorm}
\end{align}

<< "example2" >>=
## do reverse-Bayes calculations
alpha2 <- 0.1
za2 <- qnorm(p = 1 - alpha2/2)
sprior2 <- sqrt(so^2/(zo^2/za2^2 - 1))

## prior data conflict assessment
pbox2 <- pchisq(q = tr^2/(sr^2 + sprior2^2), df = 1, lower.tail = FALSE)
pboxFormat2 <- pFormat(p = pbox2)
@

Figure~\ref{fig:RBillustration} shows the prior predictive $p$-values
from~\eqref{eq:pboxNorm} computed for three example replication studies from the
``Replication Project: Cancer Biology''. We see that larger effect estimates
show smaller prior predictive $p$-values $p_{\text{Box}}$, indicating more
prior-data conflict. This is because the standard errors from all three
replications are roughly the same size, so that the distance between zero and
the replication effect estimate matters most. The $p$-values suggest that there
is hardly any conflict between the sceptical prior and the first two
replications (those with $p_{\text{Box}} = \Sexpr{pboxFormat[1]}$ and
$p_{\text{Box}} = \Sexpr{pboxFormat[2]}$), while the conflict is larger for the
third one (the one with $p_{\text{Box}} = \Sexpr{pboxFormat[3]}$).

% However, it is not clear how numerical values of $p_{\text{Box}}$ should be
% interpreted and below which value a replication should be declared
% ``successful''.
\citet{Held2020} defined replication success at level $\alpha$ by
\begin{align*}
  p_{\text{Box}} \leq \alpha.
\end{align*}
In words, replication success is established if there is more conflict between
the sceptical prior and the replication data than there was evidence against the
null hypothesis in the original study. For the examples in
Figure~\ref{fig:RBillustration}, all prior predictive $p$-values are larger than
the level $\alpha = 5\%$ used for computing the sufficiently sceptical prior
variance ($\tau^{2}_{\alpha} = \Sexpr{round(sprior, 2)}^{2}$), so neither of
them achieves replication success at level $\alpha = 5\%$. However, at a larger
level, \eg{} \mbox{$\alpha = \Sexpr{round(alpha2*100, 2)}\%$}, the corresponding
sufficiently sceptical prior variance would be smaller
($\tau^{2}_{\alpha} = \Sexpr{round(sprior2, 2)}^{2}$). Consequently, there would
be more conflict between the prior and the replication data,
% (in the same order as in Figure~\ref{fig:RBillustration}:
% $p_{\text{Box}} = \Sexpr{pboxFormat2[1]}$,
% $p_{\text{Box}} = \Sexpr{pboxFormat2[2]}$, and
% $p_{\text{Box}} = \Sexpr{pboxFormat2[3]}$),
so that the third replication would be successful at the less convincing level
\mbox{$\alpha = \Sexpr{round(alpha2*100, 2)}\%$} (since the prior predictive
$p$-value would be
$p_{\text{Box}} = \Sexpr{pboxFormat2[3]} < \Sexpr{round(alpha2*100, 2)}\%$).

<< "example3" >>=
## do reverse-Bayes calculations
alpha2 <- 0.1
za2 <- qnorm(p = 1 - alpha2/2)
sprior2 <- sqrt(so^2/(zo^2/za2^2 - 1))

## prior data conflict assessment
pbox2 <- pchisq(q = tr^2/(sr^2 + sprior2^2), df = 1, lower.tail = FALSE)
pboxFormat2 <- pFormat(p = pbox2)

zr <- tr/sr
c <- so^2/sr^2
ps <- pSceptical(zo = zo, zr = zr, c = c, alternative = "two.sided", type = "nominal")
psFormat <- pFormat(p = ps)
@
\subsubsection{The sceptical \textit{p}-value}
To remove the dependence on the level $\alpha$, \citet{Held2020} proposed to
determine the smallest level at which replication success can be established.
This level is called the \emph{sceptical $p$-value} $p_{\text{S}}$, and it is
available in closed-form
\begin{align}
  \label{eq:ps}
  p_{\text{S}}
  % = \inf_{\alpha \in (0, 1)} \left\{\alpha : p_{\text{Box}} \leq \alpha \right\}.
  = 2\left\{1 - \Phi(|z_{\text{S}}|)\right\}
\end{align}
where
\begin{align*}
  z_{\text{S}}^{2}
  = \begin{cases}
    z^{2}_{\text{H}}/2 & \text{for} ~ c = 1 \\
    \left\{\left[z_{\text{A}}^{2}\left\{z_{\text{A}}^{2} +
          z_{\text{H}}^{2}(c - 1)\right\}\right]^{1/2}
      - z_{\text{A}}^{2}\right\}/(c - 1)
    & \text{for} ~ c \neq 1
  \end{cases}
\end{align*}
with arithmetic mean $z_{\text{A}}^{2} = (z^{2}_{o} + z^{2}_{r})/2$ and harmonic
mean $z_{\text{H}}^{2} = 2/(1/z^{2}_{o} + 1/z^{2}_{r})$ of the squared
$z$-statistics, and variance ratio $c = \sigma^{2}_{o}/\sigma^{2}_{r}$.
Replication success at level $\alpha$ is then equivalent to
$p_{\text{S}} \leq \alpha$. For instance, the sceptical $p$-values of the three
replication studies in Figure~\ref{fig:RBillustration} are
$p_{\text{S}} = \Sexpr{psFormat[1]}$, $p_{\text{S}} = \Sexpr{psFormat[2]}$, and
$p_{\text{S}} = \Sexpr{psFormat[3]}$ (from left to right), so we can see that
the third replication is unsuccessful at level $\alpha = 5\%$ but successful at
level $\alpha = 10\%$. However, the sceptical $p$-value does not necessarily
have to be dichotomized in this manner, but can also be interpreted as a
quantitative measure of replication success; the smaller $p_{\text{S}}$, the
higher the degree of replication success.

The sceptical $p$-value has several interesting properties \citep[Section
3]{Held2020}: First, it is always larger than the maximum of the original and
replication $p$-values ($p_{\text{S}} > \max\{p_{o}, p_{r}\}$), meaning that
both $p$-values have to be smaller than $\alpha$ such that replication success
at level $\alpha$ is possible. The sceptical $p$-value hence requires
\emph{both} studies to be sufficiently convincing on their own (in terms of
their $p$-values), similar to the significance criterion for replication
success. Second, if the $p$-values $p_{o}$ and $p_{r}$ remain fixed but the
relative effect estimate \mbox{$d = \that_{r}/\that_{o}$} decreases towards
zero, the sceptical $p$-value increases, meaning that shrinkage of the
replication effect estimate is penalized ($p_{\text{S}} \uparrow 1$ as
$d \downarrow 0$ for fixed $p_{o}$ and $p_{r}$). The sceptical $p$-value thus
takes effect shrinkage into account, unlike the significance criterion which can
be achieved with any non-zero replication effect estimate $\that_{r}$ provided
the standard error $\sigma_{r}$ is small enough. This property is desirable in
the replication setting as a smaller replication effect estimate $\that_{r}$ may
not be practically relevant anymore, despite its statistical significance.

There is one problem with the (two-sided) sceptical $p$-value as defined
in~\eqref{eq:ps}; It does not take the direction of the effect estimates into
account, so replication success may be established even though the effect
estimate from the replication goes in the opposite direction of the original one
($\sign(\that_{o}) \neq \sign(\that_{r})$). In fact, the sceptical $p$-value
will decrease with decreasing negative relative effect estimate
($\ps \downarrow p_{o}$ as $d \downarrow -\infty$ for fixed $\sigma_{o}$ and
$\sigma_{r}$) which seems contrary to intuitive understandings of replicability
as a change in effect direction may diminish the usefulness of a scientific
finding (\eg{} a negative treatment effect may indicate harm instead of
benefit).
% For example, if $\that_{o} = \Sexpr{smdo}$,
% $\that_{r} = \Sexpr{smdr}$, and $\sigma_{o} = \sigma_{r} = 1$, this would lead
% to $\ps = \Sexpr{pFormat(ps1)}$, and the sceptical $p$-value continues to
% decreases with decreasing replication effect estimates.
This ``replication
paradox'' \citep{Ly2018} can be avoided by defining a one-sided sceptical
$p$-value
\begin{align*}
  p_{\text{S}} =
  \begin{cases}
   1 - \Phi(|z_{\text{S}}|) & \text{if} ~\sign(\that_{o}) = \sign(\that_{r}) \\
   \Phi(|z_{\text{S}}|) & \text{if} ~ \sign(\that_{o}) \neq \sign(\that_{r}) \\
  \end{cases}
\end{align*}
similar to how ordinary one-sided $p$-values are defined. The one-sided
sceptical $p$-value has similar properties as the two-sided version -- it is
always larger than the one-sided $p$-values from original and replication
studies, and it penalizes shrinkage of the replication effect estimate. In the
subsequent chapters only the one-sided version of the sceptical $p$-value will
be used.

The last column of Table~\ref{tab:RPCB} shows the one-sided sceptical $p$-value
for the subset of replications from the ``Reproducibility Project: Cancer
Biology''. The results illustrate the previously discussed properties to require
significance from both studies and to penalize effect shrinkage. For example,
replication \#4 fails to achieve replication success at level $\alpha = 2.5\%$
with the one-sided sceptical $p$-value, even though the replication study was
highly convincing (the effect estimate was almost five times as large as in the
original study). Yet, as the approach requires both studies to be convincing on
their own -- and the original study was not significant at the 2.5\% level
(one-sided) -- replication success at this level is impossible with the
sceptical $p$-value. The second property is illustrated by replication \#8. Here
both the original and replication studies were significance. However, the effect
estimate from the replication was roughly 60\% smaller than the one from the
original study, and significance is merely achieved because the standard error
was much smaller (\ie{} around $1/\surd c \approx 1/2.8$ times smaller). The
sceptical $p$-value is therefore only $p_{\text{S}} = 0.12$, indicating less
evidence for replication success.

<< "paradox-example" >>=
smdo <- 3
smdr <- -3
so <- sr <- 1
ps1 <- pSceptical(zo = smdo/so, zr = smdr/sr, c = so^2/sr^2,
                  alternative = "two.sided", type = "nominal")
@

This concludes the introduction to replication studies and reverse-Bayes methods
for their analysis. Some additional properties (\eg{} the null distribution of
the sceptical $p$-value) and extensions (\eg{} power and sample size
calculations) can be found in the original article by \citet{Held2020}. The
sceptical $p$-value and related methods are implemented in the R package
\texttt{ReplicationSuccess} available on CRAN
(\url{https://cran.r-project.org/package=ReplicationSuccess}). % However, the
% relevant parts of the procedure have been presented.
In the following, I will discuss open questions and how they are addressed in
this thesis.

\section{Thesis contributions}
\label{sec:summary}
\raggedbottom

This thesis consists of six papers divided into three parts. The first part
(Paper~\ref{PaperI}, \ref{PaperII}, and \ref{PaperIII}) focuses on methodology
for design and analysis of replication studies. The second part
(Paper~\ref{PaperIV} and \ref{PaperV}) revolves around reverse-Bayes methodology
with a broader scope of applications than replication
studies. % Paper IV is a review paper about
% reverse-Bayes methodology and Paper V is a short comment on another article
% which proposed a reverse-Bayes method.
The last part (Paper~\ref{PaperVI}) deals with integrity issues in
methodological research.
% The article lists and illustrates questionable
% research practices in simulation studies, and gives recommendations how to avoid
% them.

\subsection{Design and analysis of replication studies}
% The research from this parts extends

\startPaperSummary{PaperI}{%
  The assessment of replication success based on relative effect size}{%
  Leonhard Held, Charlotte Micheloud, Samuel Pawel \\
  \textit{The Annals of Applied Statistics}, 2022, 16(2), 706--720.
  doi:\href{https://doi.org/10.1214/21-AOAS1502}{10.1214/21-AOAS1502}.}

It is not clear how to interpret the sceptical $p$-value as it is not an
ordinary $p$-value (which has a uniform distribution under the null hypothesis
that the underlying effect size $\theta$ is zero). Moreover, when the same level
$\alpha$ as for the significance criterion is used for dichotomizing the
sceptical $p$-value, this leads to a much more stringent criterion for
replication success than the significance criterion. Many studies which
intuitively seem successful will not satisfy it, and frequentist properties such
as power and type I error rate will be impacted by it (\eg{} power and type I
error rate will be much lower compared to the significance criterion).

<< "golden-level" >>=
alpha <- 0.025
alphaG <- levelSceptical(level = alpha, alternative = "one.sided", type = "golden")
@

In this article, we therefore look closer at the ``success region'' of the
sceptical $p$-value in terms of the relative effect estimate
$d = \that_{r}/\that_{o}$. This perspective leads to the proposal of a new
default level for thresholding the sceptical $p$-value called the \emph{golden
  level} $\alpha_{G}$ (named ``golden'' because the golden ratio appears in its
derivation). The golden level is defined through the property that for an
original study which was borderline significant ($p_{o} = \alpha$), replication
success based on $p_{\text{S}} \leq \alpha_{G}$ is only possible if the
replication effect estimate is at least as large as the original one
($d \geq 1$). For instance, for the one-sided significance level
$\alpha = \Sexpr{round(alpha*100, 1)}\%$, the corresponding two-sided golden
level is $\alpha_{G} = \Sexpr{round(alphaG*100, 1)}\%$. The behavior of the
golden level seems to align with common sense: for original studies which were
already convincing (in terms of their $p$-value) the effect estimate in the
replication study is allowed to shrink, to some extent, whereas for less
convincing original studies (those with $p$-values around the significance
level) shrinkage is more strongly penalized. We find that for replication
studies with equal or larger sample size than the original study, replication
success based on the golden level controls the type I error rate at the same
level as the standard significance criterion. Similarly, if the original study
was adequately powered, replication success based on the golden level has
comparable or improved ``project power''compared to the significance criterion
(project power is the probability of replication success conditional on a true
effect size based on original and replication study). Case studies from four
large-scale replication projects \citep{Opensc2015, Camerer2016, Camerer2018,
  Cova2018} illustrate the properties of the golden level. In most cases
conclusions agree with the significance criterion, but in the differing cases,
the golden level seems to produce more appropriate inferences as it takes into
account effect shrinkage. This extension is now implemented as the default
option in the R package \texttt{ReplicationSuccess}.

L. Held had the idea to apply the sceptical $p$-value to the data from the four
replication projects, which I collected for my master thesis. L. Held and C.
Micheloud came up with the golden level. L. Held wrote an initial draft of the
manuscript. L. Held, C. Micheloud, and I then iteratively worked on the
manuscript.

\startPaperSummary{PaperII}{%
  The sceptical Bayes factor for the assessment of replication success}{%
  Samuel Pawel, Leonhard Held \\
  \textit{Journal of the Royal Statistical Society: Series B (Statistical
    Methodology)}, 2022, 84(3), 879--911.
  doi:\href{https://doi.org/10.1111/rssb.12491}{10.1111/rssb.12491}.}

The reverse-Bayes approach from \citet{Held2020} is based on challenging the
original study with a sceptical prior so there is no longer evidence for an
effect. Evidence is quantified in terms of credible intervals and tail
probabilities. However, there exist also other measures of evidence, and it has
been a matter of long debates which is the most appropriate \citep[see
\eg{}][]{Berger1987, Casella1987, Royall1997, Berger2003, Benjamin2017,
  Lakens2018, Amrhein2019b}. In this paper, we extend the reverse-Bayes
assessment of replication studies to use Bayes factors \citep{Good1958,
  Jeffreys1961} for quantifying evidence and prior-data
conflict. % Bayes factors are also
% used for prior-data conflict assessment as suggested by \citep{Box1980}.
Similar to the sceptical $p$-value, the procedure leads to a single measure for
quantifying replication success, \emph{the sceptical Bayes factor}. Systematic
comparisons show that the sceptical Bayes factor has similar properties as the
sceptical $p$-value %due to the reverse-Bayes approach underlying both methods
(\eg{} it penalizes shrinkage of the replication effect estimate and requires
compelling evidence from both studies in terms of $p$-values and Bayes factors,
respectively). However, it is also shown that the sceptical $p$-value suffers
from a certain type of ``shrinkage paradox'', which is avoided by the sceptical
Bayes factor; when the $p$-value from the original study goes to zero,
replication success based on the sceptical $p$-value can be achieved with any
arbitrarily small replication effect estimate, whereas replication success based
on the sceptical Bayes factor poses a finite limit on how much shrinkage is
allowed. Technically, the procedure is more involved and closed-form solutions
for the sceptical Bayes factor are only available in special
situations. % and in terms of the
% ``Lambert W function'' \citep{Corless1996}.
The method is illustrated on data from the ``Social Sciences Replication
Project'' \citep{Camerer2018}, and implemented in the R package
\texttt{BayesRep} (\url{https://gitlab.uzh.ch/samuel.pawel/BayesRep}).

The idea to use Bayes factors instead of tail probabilities was suggested by
\citet{Consonni2019} and \citet{Pericchi2020} independently in response to the
original article by \citet{Held2020}. L. Held then implemented a first version
of the procedure for the grant application of this research project
\citep{Heldproposal2019}. I then worked out the technical and implementation
details, including closed-form solution for the sceptical Bayes factors,
asymptotic properties, type I and type II error rates, and non-normal
extensions. I wrote the initial draft of the manuscript and the R package.
Throughout, L. Held gave high-level feedback. I presented initial results at the
annual meeting of the GMDS and CEN-IBS (German Association for Medical
Informatics, Biometry and Epidemiology and the Central European Network of the
International Biometric Society) in 2020. L. Held presented the final results at
the ISBA (International Society for Bayesian Analysis) world meeting 2021.

\startPaperSummary{PaperIII}{%
  Bayesian approaches to designing replication studies}{%
  Samuel Pawel, Guido Consonni, Leonhard Held \\
  \textit{arXiv preprint}, 2022.
  doi:\href{https://doi.org/10.48550/arXiv.2211.02552}{10.48550/arXiv.2211.02552}.}

An important aspect in the design of replication studies is determining their
sample size. How exactly the sample size should be determined depends on which
method is used for the analysis of the replication data. Various approaches have
been proposed for doing so which are specifically tailored to certain analysis
methods. In this article, we provide a general Bayesian framework which applies
to any analysis method (including the sceptical $p$-value and the sceptical
Bayes factor). We show how the data from the original study and external
knowledge can be combined in a \emph{design prior} for the underlying model
parameters. Based on this design prior, predictions about the replication data
can be computed, and the replication sample size can be chosen such that the
probability of replication success becomes as high as desired. We illustrate
Bayesian design of replication studies in the normal-normal hierarchical model
which provides sufficient flexibility for specification of design priors. Data
from a cross-laboratory replication project \citep{Protzko2020} are used for
illustrating our methods, which are available in the R package
\texttt{BayesRepDesign} (\url{https://github.com/SamCH93/BayesRepDesign}).

L. Held specified in the grant application of this research project
\citep{Heldproposal2019} that we will investigate power and sample size
calculations for the sceptical $p$-value and the sceptical Bayes factor. In
Paper~\ref{PaperII}, I already derived the power function of the sceptical Bayes factor in
closed-form for two types of design priors. After its completion, I generalized
the result to any design prior in the normal-normal hierarchical model, and
started working on this manuscript. I presented a first draft to L. Held and G.
Consonni in the beginning of 2021. \mbox{G. Consonni} then helped developing the
methodology for multisite replication study design. I continued working on the
manuscript in 2022 and also wrote the accompanying R package. Throughout, L.
Held and G. Consonni gave high-level feedback.

\subsection{Reverse-Bayes methodology}

\startPaperSummary{PaperIV}{%
  Reverse-Bayes methods for evidence assessment and research synthesis}{%
  Leonhard Held, Robert Matthews, Manuela Ott, Samuel Pawel \\
  \textit{Research Synthesis Methods}, 2022, 13(3), 295--314.
   doi:\href{https://doi.org/10.1002/jrsm.1538}{10.1002/jrsm.1538}.}

While the popularity of Bayesian methods has been rapidly increasing since the
advent of modern computational methods in the 1990s, reverse-Bayes methods % , such
% as the reverse-Bayes assessment of replication studies,
have remained largely unknown to statisticians and users of statistics alike. In
this article, we review reverse-Bayes history and methods to increase awareness
about the approach. Specifically, we summarize the work on reverse-Bayes by I.
J. Good \citep{Good1950}, who first proposed the idea. We then review methods
such as the \emph{Analysis of Credibility} from \citet{Matthews2001b,
  Matthews2018}, its extension to Bayes factors, and the \emph{False Positive
  Risk} from \citet{Colquhoun2017}. To illustrate these method, we use data from
a meta-analysis on the effect of corticosteroids on COVID-19 mortality.

L. Held and M. Ott started working on this article several years ago. When I
discovered the connection between the Analysis of Credibility and the fail-safe
$N$ method \citep{Rosenthal1979,Rosenberg2005}, L. Held suggested that it would
fit nicely into this manuscript and to add the COVID-19 meta-analysis example. I
rewrote and expanded his initial draft, adding also a new section on
reverse-Bayes approaches with Bayes factors, largely based on the work from
Paper~\ref{PaperII}. We then managed to recruit R. Matthews to also contribute.
From that point on the three of us iteratively worked on the manuscript and M.
Ott gave high-level feedback.

\startPaperSummary{PaperV}{%
  Comment on ``Bayesian additional evidence for decision making under small
sample uncertainty''}{%
   Samuel Pawel, Leonhard Held, Robert Matthews \\
  \textit{BMC Medical Research Methodology}, 2022, 22(149).
  doi:\href{https://doi.org/10.1186/s12874-022-01635-4}{10.1186/s12874-022-01635-4}.}

Shortly after the acceptance of Paper~\ref{PaperIV}, the article by
\citet{Sondhi2021} appeared. It proposed a novel reverse-Bayes method called
\emph{Bayesian Additional Evidence}, and we noted some flaws in the article.
This prompted us to write a commentary. We show that -- contrary to the
statement by Sondhi et al. -- there is a closed form solution for the key
quantity in their approach termed ``Bayesian Additional Evidence tipping
point''.
% The Bayesian Additional Evidence method provides an answer to the question
% ``What prior do I have to choose so that that my unconvincing data become
% convincing?''.
The method is also closely related to the Analysis of Credibility by
\citet{Matthews2018}. We investigate differences and similarities of the two
methods, concluding that the priors determined through the Bayesian Additional
Evidence method are of limited use due to their restrictive assumption of having
equal variance as the data.

R. Matthews alerted us about the article from \citet{Sondhi2021}. After reading
it, I realized that their statement about closed-form solutions was incorrect
and derived a solution. \mbox{L. Held} suggested to write a commentary. I wrote
an initial draft of the manuscript, which \mbox{R. Matthews} improved by
discussing some logical flaws of the method. The two of us iteratively worked on
the manuscript, while L. Held gave high-level feedback.

\subsection{Meta-scientific perspectives on methodological research}

\startPaperSummary{PaperVI}{%
  Pitfalls and Potentials in Simulation Studies}{%
  Samuel Pawel, Lucas Kook, Kelly Reeve \\
  \textit{arXiv preprint}, 2022.
  doi:\href{https://doi.org/10.48550/arXiv.2203.13076}{10.48550/arXiv.2203.13076}.}

In methodological research, simulation studies are frequently used for
evaluating ``how well'' statistical methods perform. The conclusions from these
studies are therefore often crucial for recommendations when which method should
be used. However, several literature review articles identified that the design
and reporting standards of simulation studies have remained low over the years
\citep{Hoaglin1975, Burton2006, Morris2019}. Moreover, some authors have
recently argued that methodological research is suffering from reproducibility
issues, publication bias, and a ``replication crisis'' due to researchers
engaging in questionable research practices, such as selective reporting
\citep{Boulesteix2020}. In this article, draw attention to these issues. We
summarize possible questionable research practices in simulation studies, and
show how easy it is to make a method seem superior if various questionable
research practices are employed. We also give recommendations to alleviate these
issues. Most importantly, we recommend researchers to write and pre-register
simulation protocols.

The first authorship of this manuscript is shared with L. Kook. I had the idea
to invent a novel method and use questionable research practices that make it
seem superior, to draw attention to the low standards in methodological
research. I then wrote a first draft of the manuscript and proposed the idea to
L. Kook and K. Reeve. L. Kook and myself then invented the regression method
``AINET'' and started writing the simulation protocol. L. Kook took lead in
developing the R package \texttt{ainet}
(\url{https://github.com/LucasKook/ainet}) and simulation study code. Myself, L.
Kook, and K. Reeve then designed the final simulation study and continued
working on the manuscript together. Recently, I was invited to present the
results from this project at the CEN (Central European Network) conference 2023
in Basel.

\phantomsection
\addcontentsline{toc}{section}{Data and software}
\section*{Data and software}
The CC-By 4.0 licensed data from the ``Reproducibility Project: Cancer Biology''
\citep{Errington2021} were downloaded from
\url{https://doi.org/10.17605/osf.io/e5nvr}. The relevant variables were then
extracted from the file ``\texttt{RP\_CB Final Analysis - Effect level
  data.csv}''. All analyses from this chapter were conducted in the R
programming language version \Sexpr{paste(version$major, version$minor, sep =
  ".")} \citep{R}. The packages \texttt{dplyr} \citep{Wickham2022},
\texttt{ggplot2} \citep{Wickham2016}, \texttt{knitr} \citep{Xie2022},
\texttt{ReplicationSuccess} \citep{Held2020}, \texttt{UpSetR}
\citep{Gehlenborg2019}, and \texttt{xtable} \citep{Dahl2019} were used. Code and
data to reproduce this thesis are available at
\url{https://github.com/SamCH93/thesis}. A snapshot of the git repository is
archived at \url{https://doi.org/10.5281/zenodo.XXXXXXX}.

<< "sessionInfo1", eval = Reproducibility, results = "asis" >>=
## print R sessionInfo to see system information and package versions
## used to compile the manuscript (set Reproducibility = FALSE, to not do that)
cat("\\newpage \\section*{Computational details}")
@
<< "sessionInfo2", echo = Reproducibility, results = Reproducibility >>=
sessionInfo()
@

% Bibliography
\nocite{Pawel2020b, Held2021, Pawel2022c,Held2021b, Pawel2022d, Pawel2022e}
\bibliographystyle{apalikedoiurl}
\bibliography{bibliography}